### 2016-07-29 Doc/MSelection/mselection.tex
mselection stuff

% Template for PLoS
%subsubsection is 3rd level heading, PLOS doen't support 4th level.
% Version 3.3 June 2016
% Please follow the template instructions whenever possible.
%
% Once your paper is accepted REMOVE ALL TRACKED CHANGES.
% Use latexdiff to track changes during review to maintain a clean file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
% Do not include colors or graphics in the text.
%
% LaTeX source should be a single file: don't use \input, \externaldocument
%
% -- FIGURES AND TABLES
%
% Include table/figure captions directly after paragraph where first cited.
%
% NO GRAPHICS
% - Figures should be uploaded separately from your manuscript file. 
% - Extract figures (in PDF) generated by LaTeX before submission.
% - Combine multiple panels/subfigures into one image before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}
%\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
%\rfoot{\thepage/\pageref{lastpage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Mutation Rate as a Selection Mechanism in Continuous Evolution of Bacteriophage} % Title Case
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Peter Reintjes\textsuperscript{1}
\\
\bigskip
\textbf{1} Museum of Life and Science, Durham, North Carolina, USA
\bigskip

% Additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
%* correspondingauthor@institute.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Engineering novel proteins via continuous evolution of bacteriophage currently requires the host bacteria to be transformed with two additional functions: A mutagenesis vector to provide an elevated rate of viral mutation, and a selection mechanism to increase the number of infectious progeny for the mutant encoding the improved product\cite{pace}.  I propose using a selection mechanism which simply modulates the rate of mutation by producing a mutagenesis suppressor from the desired activity of the evolving protein.  Improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.

As the evolving protein approaches a high level of desired activity, the proportionally lowered mutation rate will ensure that more viral progeny contain exact copies of the parent genotype while a baseline mutation rate ensures that our mechanism to lower the mutation rate will not result in stagnation.

Extremely high in vivo mutation rates are now possible\cite{mutation}.  With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction. The modulation of mutation rate within each host cell means that one of these wildly mutating strains could find an environment for faithful reproduction if it should stumble upon a promising mutation in the evolving gene.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
%\section*{Author Summary}
%Lorem ipsum dolor sit amet,

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
\chapter{M-Selection}
\section{Introduction}
Continuous evolution of bacteriophage has become a potent protein engineering
tool.  For example, PACE - Phage Assisted Continuous Evolution described by
Esvelt,Carlson and Lui.

The continuous evolution process produces a viral genome containing a gene
which has undergone many generations of mutation and selective pressure for
a particular property.  The generality of this approach to protein engineering
is only limited by our ability to create a selection mechanism for the desired
property and insert an expressable initial form of the gene into the virus.

Specifically, the PACE system requires:
\begin{itemize}

\item{}
A modified viral genome missing a gene required for infectious progeny.

\item{}
A modified host genome/plasmid to induce an elevated mutation rate.

\item{}
A modified host genome/plasmid to provide a selection mechanism to provide
the missing gene product in proportion to the desired activity of the evolving
gene.
\end{itemize}

Effectively, the number of infectious progeny is proportional to the activity
of the selection mechanism. In principle, if the current mutation of the gene
produces a product with none of the desired activity, none of the infectivity
protein will be produced by the host, and there will be no infectous progeny.
This process involves a continuous flow of host cells through a reaction vessel
and so any genotype producing less than two infectious virions per host will
be washed out.

M-Selection is proposed to the selection mechanism by combining it with
the mechanism for elevated mutation.

The benefits of M-Selection may include:

The virus is not crippled, but only has an additional gene or domain, similar to
phage display.

The nominal (high) mutation rate can be chosen to guarantee that there will be
a minimum number of non-mutant copies produced for every infection.  This percentage of clones will then increase as the mutation rate is reduced.

For example, if the selection mechanism stimulated by the desired activity will consist
of a repressor for the production of mutagenic polymerase II-.
Positive selection is the result of reduced mutation, resulting in more clones of the
selected virus relative to its mutants.

As far as we know, modulation of the mutation rate within a particular host has
not been exploited in this way.  This form of albeit un-natural selection is referred to
as M-selection, whereby the selection mechanism consists of reducing the rate of mutation
acting on a particular instance of the viral genome precisely because it is having some
desired behavior.

Figure 1 shows how the combined mutagenic and selection plasmid takes the place of the AP and MP in PACE.

\def\boxone#1{
  \put(19,30){\line(1,0){67}}     % horiz top
  \put(19,0){\line(1,0){67}}      % horiz bottom
  \put(20,0){\line(0,1){30}}      % vert left
  \put(85,0){\line(0,1){30}}      % vert right
  \put(25,15){#1}
}
\def\boxtwo#1#2{
  \put(19,30){\line(1,0){80}}     % horiz top
  \put(19,0){\line(1,0){80}}      % horiz bottom
  \put(20,0){\line(0,1){30}}      % vert left
  \put(98,0){\line(0,1){30}}      % vert right
  \put(25,17){#1}
  \put(25,7){#2}
}
\def\circleone#1#2{\circle{#1}\put(-17,-4){#2}}
\def\circletwo#1#2#3{\circle{#1}\put(-17,0){#2}\put(-15,20){#3}}

\picture(200,150)(-70,0)

\put(90,90){\vector(0,-1){15}}
\put(90,30){\vector(0,1){15}}

\put(20,60){\vector(1,0){30}}
\put(130,60){\vector(1,0){30}}
\put(130,60){\vector(1,1){30}}
\put(130,60){\vector(1,-1){30}}
\linethickness{2pt}
\thicklines
\put(0,60){\circleone{40}{\Large{\ \ P1}}}
\put(30,90){\boxtwo{\ Phage Shock}{\ \ \ Promoter}}
\put(30,45){\boxtwo{MUTAGENESIS}{\ \ Replication}}
\put(30,0){\boxtwo{Antisense RNA}{\ \ Suppressor}}

\put(180,60){\circleone{40}{\Large{\ \ F1}}}

\endpicture
\vskip 1.0cm
\centerline{\bf Fig 1. M-selection Plasmid}
\vskip 0.2cm

\subsection{Difficulties of working with crippled Virus genome}

\begin{eqnarray}
\label{eq:schemeP}
	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
\end{eqnarray}

\subsection*{Multiplicity of Infection}

\vskip 0.2cm
\begin{eqnarray}
\label{eq:MOI}
\mathrm{P}\left( x \right) = \frac{{e^{ - \lambda } \lambda ^x }}{{x!}}
\end{eqnarray}

\section{Comparison with Natural Selection}
Selection pressure for an optimal mutation rate has its own natural history\cite{evomut} although for settled 
A selection mechanism for lowered mutation rate
Natural selection is largely about responses to the external environment, which we control closely in laboratory evolution.
The closest parallel would be found in species which can reproduce by asexual or sexual reproduction and change the rates between the alternatives as a way of accessing more diversity through sexual reproduction or producing more clones via asexual reproduction.

\section*{Materials and Methods}
\subsection*{Components of Statistical Analysis}
Percentage of activated genes due to Phage Shock promoter.
Number of polII-  (mutagenic polymerase) transcripts * error rate due to mutagenic polymerase.

Error rate of normal polII
Error rate of polII-

Number of phage particles produced np(0-6 min) = 0 (six minutes before any phage appear),
then production between 75-200 per hour.

Initial single-strand (virion) DNA production occurs in non-mutagenic environment, then mutants appear along with polII- production, until suppression of polII- occurs due to binding.  Binding ? anti-sense RNA transcription ? polII-  repression ? reduction in mutations.

% For figure citations, please use "Fig" instead of "Figure".
Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi.

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}
\subsubsection*{3rd Level Heading.} 
Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\begin{enumerate}
	\item{react}
	\item{diffuse free particles}
	\item{increment time by dt and go to 1}
\end{enumerate}

\subsection*{Sed ac quam id nisi malesuada congue.}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{itemize}
	\item First bulleted item.
	\item Second bulleted item.
	\item Third bulleted item.
\end{itemize}

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.

\section*{Conclusion}

I think this will work.

\section*{Supporting Information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem Ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem Ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
% 
\begin{thebibliography}{10}

\bibitem{pace}
Esvelt K. M., Carlson J. C. \& Liu D. R.
\newblock {A system for the continuous directed evolution of biomolecules}.
\newblock Nature 472, 499–503 (2011) .

\bibitem{mutation}
Badran AH, Liu DR.
\newblock {{D}evelopment of potent in vivo mutagenesis plasmids with broad mutational spectra}.
\newblock Nature Communications. 2015;6:8425. doi:10.1038/ncomms9425.

\bibitem{evomut}
Lynch M.
\newblock {{E}volution of the Mutation Rate}.
\newblock Trends in Genetics, 2010 August ; 26(8): 345–352. doi:10.1016/j.tig.2010.05.003.
\end{thebibliography}

\end{document}

### 2016-07-29 Doc/MSelection/mselection.tex
remove lorem

@@
 \section{Comparison with Natural Selection}
-Selection pressure for an optimal mutation rate has its own natural history\cite{evomut} although for settled 
-A selection mechanism for lowered mutation rate
+Selection pressure for an optimal mutation rate has its own natural history\cite{evomut} although for settled phyla we see a certain natural mutation rates.
+Bacteriophage and multicellular eukaryote mutation rates are higher than bacteria and single cell eukaryotes (around 0.003).
+
+A selection mechanism for lowered mutation rate would not seem to correspond to anything in the natural world.
 Natural selection is largely about responses to the external environment, which we control closely in laboratory evolution.
-The closest parallel would be found in species which can reproduce by asexual or sexual reproduction and change the rates between the alternatives as a way of accessing more diversity through sexual reproduction or producing more clones via asexual reproduction.
+Perhaps the closest parallel is found in species which can reproduce by asexual or sexual reproduction and change the rates between the alternatives as a way of accessing more diversity through sexual reproduction or producing more clones via asexual reproduction.
 
 \section*{Materials and Methods}
-\subsection*{Components of Statistical Analysis}
+
+\subsection*{Population Dyamics of Bacteriophage}
+\subsubsection*{Clones verses Mutants}
+Although it may sounds like the title of a recent movie, this precisely states the selection mechanism of of mutation.
+
 Percentage of activated genes due to Phage Shock promoter.
-Number of polII-  (mutagenic polymerase) transcripts * error rate due to mutagenic polymerase.
+Number of polIII-  (mutagenic polymerase) transcripts times error rate due to mutagenic polymerase.
 
-Error rate of normal polII
-Error rate of polII-
+Error rate of normal polII, polIII
 
@@
 
-% For figure citations, please use "Fig" instead of "Figure".
-Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius.
-
-% Place figure captions after the first paragraph in which they are cited.
-\begin{figure}[!h]
-\caption{{\bf Bold the figure title.}
-Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-\label{fig1}
-\end{figure}
-
 % Results and Discussion can be combined.
 \section*{Results}
-Nulla mi mi.
 
-% Place tables after the first paragraph in which they are cited.
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-\centering
-\caption{
-{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
-\begin{tabular}{|l+l|l|l|l|l|l|l|}
-\hline
-\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
-$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
-$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
-$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
-\end{tabular}
-\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-\end{flushleft}
-\label{table1}
-\end{adjustwidth}
-\end{table}
-
-\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}
-\subsubsection*{3rd Level Heading.} 
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
-Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
+\subsection*{Comparison with Sexual/Asexual Reproduction}
+Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copied of its genome.
+For organisms with access to both forms, the percentage of offspring produced by asexual reproduction would be comparable to our percentage of true copies.
+\subsubsection*{Yeast}
+Although yeast has access to both sexual and asexual reproduction, the (time) cost of sexual reproduction is high and it is generally invoked under conditions of stress, where an increase in genetic variety of offspring.
+\subsubsection*{Daphnia} 
+In Daphnia, longevity and healthy survival to the end of the season result in so called 'winter eggs' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
+The cost differential between sexual and asexual reproduction is low and each generation represents a mix of both.
 
@@
 
-\subsection*{Sed ac quam id nisi malesuada congue.}
+\subsection*{Using RNA Antisense.}
 
-Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
+The components for the broad spectrum, high level mutagenisis described in Badran and Liu\cite{mutation} are variations (often single mutation variants) of the functional domains of high fidelity DNA replication and as such are not suitable targets for RNA anti-sense interference.
+Any such anti-sense interference would be likely to affect their normal counterparts which are needed for faithful reproduction.
+
+RNA anti-sense of gIV could be utilized to block the primary phage shock promotor which was the primary trigger activating the high level of mutation in the lagoon.
+
+Lowering the mutation rate can be achieved in a number of ways.
 
 \begin{itemize}
-	\item First bulleted item.
-	\item Second bulleted item.
-	\item Third bulleted item.
+	\item Supress pIV production or otherwise block phage shock promotor
+	\item Selectively block downstream components of mutagenesis
+	\item Enhance production of wild-type high-fidelity replication components.
 \end{itemize}
@@
 \section*{Discussion}
-Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero~\cite{bib3}.
+Modeling needs work.
 
@@
 
-\paragraph*{S2 Fig.}
-\label{S2_Fig}
-{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 File.}
-\label{S1_File}
-{\bf Lorem Ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 Video.}
-\label{S1_Video}
-{\bf Lorem Ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
 \paragraph*{S1 Appendix.}
 \label{S1_Appendix}
-{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+{\bf Algorithm.} The code goes here.
 
@@
 \label{S1_Table}
-{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+{\bf Graph maybe.} The Graph goes here.
 
 \section*{Acknowledgments}
-Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nam id pretium nisi.
-Sed ac quam id nisi malesuada congue.
-Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
+So many people to thank.
 
@@
 
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% or
-%
-% Compile your BiBTeX database using our plos2015.bst
-% style file and paste the contents of your .bbl file
-% here.
-% 
+% Alternative to directly entering bibtex entries:
+% 1) Compile BiBTeX database using plos2015.bst style file
+% 2) Paste the contents of your .bbl file here.
 \begin{thebibliography}{10}

### 2016-07-30 Doc/MSelection/mselection.tex
distr

@@
 \section*{Abstract}
-Engineering novel proteins via continuous evolution of bacteriophage currently requires the host bacteria to be transformed with two additional functions: A mutagenesis vector to provide an elevated rate of viral mutation, and a selection mechanism to increase the number of infectious progeny for the mutant encoding the improved product\cite{pace}.
-I propose using a selection mechanism which simply modulates the rate of mutation by producing a mutagenesis suppressor from the desired activity of the evolving protein.
-Improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.
-
-As the evolving protein approaches a high level of desired activity, the proportionally lowered mutation rate will ensure that more viral progeny contain exact copies of the parent genotype while a baseline mutation rate ensures that our mechanism to lower the mutation rate will not result in stagnation.
-
-Extremely high in vivo mutation rates are now possible\cite{mutation}.
-With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction.
-The modulation of mutation rate within each host cell means that one of these wildly mutating strains could find an environment for faithful reproduction if it should stumble upon a promising mutation in the evolving gene.
+Engineering novel proteins via continuous evolution of bacteriophage currently requires the host bacteria to be transformed with two additional functions: A mutagenesis vector to provide an elevated rate of viral mutation, and a selection mechanism to increase the number of infectious progeny for the individual encoding the improved product\cite{pace}.
+I propose using a mutagenesis suppressor as the selection mechanism whereby the desired activity results in a higher number of clones, thus rewarding the genotype with more faithful progeny.
 
@@
 \section{Introduction}
-Continuous evolution of bacteriophage has become a potent protein engineering
-tool.  For example, PACE - Phage Assisted Continuous Evolution described by
-Esvelt,Carlson and Lui.
-
+Continuous evolution of bacteriophage is becomming a potent protein engineering
+tool\cite{pace}\cite{monsanto}.
+For example, PACE - Phage Assisted Continuous Evolution described by Esvelt,Carlson and Lui. 
+%M13 genes
+%I(assembly),
+%II(replication init),
+%X (DNA accumulation, inside gII) 
+%III(fusion),
+%IV(assembly),
+%V(destabilizing protein),
+%VI(minor coat),
+%VII(minor coat),
+%VIII(major coat),
+%IX(minor coat),
 The continuous evolution process produces a viral genome containing a gene
@@
 \end{itemize}
+
+Extremely high in vivo mutation rates are now possible\cite{mutation}.
+With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction.
+The modulation of mutation rate within each host cell means that one of these wildly mutating strains could find an environment for faithful reproduction if it should stumble upon a promising mutation in the evolving gene.
+
+which simply modulates the rate of mutation by producing a mutagenesis suppressor from the desired activity of the evolving protein.
+Improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.
+
+As the evolving protein approaches a high level of desired activity, the proportionally lowered mutation rate will ensure that more viral progeny contain exact copies of the parent genotype while a baseline mutation rate ensures that our mechanism to lower the mutation rate will not result in stagnation.
 
@@
 
-\picture(200,150)(-70,0)
+\picture(400,150)(-90,0)
 
 \put(90,90){\vector(0,-1){15}}
-\put(90,30){\vector(0,1){15}}
+\put(90,45){\vector(0,-1){14}}
 
-\put(20,60){\vector(1,0){30}}
-\put(130,60){\vector(1,0){30}}
-\put(130,60){\vector(1,1){30}}
-\put(130,60){\vector(1,-1){30}}
 \linethickness{2pt}
 \thicklines
-\put(0,60){\circleone{40}{\Large{\ \ P1}}}
-\put(30,90){\boxtwo{\ Phage Shock}{\ \ \ Promoter}}
-\put(30,45){\boxtwo{MUTAGENESIS}{\ \ Replication}}
-\put(30,0){\boxtwo{Antisense RNA}{\ \ Suppressor}}
+\put(0,15){\circleone{40}{\Large{\ \ P1}}}
+\put(20,15){\vector(1,0){30}}
 
-\put(180,60){\circleone{40}{\Large{\ \ F1}}}
+\put(35,90){\boxtwo{\ Phage Shock}{\ \ \ Promoter}}
+\put(-20,55){
+\fbox{%
+    \begin{minipage}{1.0in}
+Activity Induced\linebreak
+\hspace*{5 mm}Repressor
+    \end{minipage}%
+}}
+\put(64,60){\vector(1,0){18}}
+\put(83,57){\fbox{\begin{minipage}{0.05in}
+		\hfill\vspace{0.30in}
+            \end{minipage}
+           }}
+\put(35,0){\boxtwo{MUTAGENESIS\ }{\ \ \ Components}}
+
+\put(130,15){\vector(1,0){30}}
+\put(180,15){\circleone{40}{\Large{\ \ F1}}}
 
@@
 \subsection*{Population Dyamics of Bacteriophage}
-\subsubsection*{Clones verses Mutants}
+\subsubsection*{Clones versus Mutants}
 Although it may sounds like the title of a recent movie, this precisely states the selection mechanism of of mutation.
@@
 \newblock Trends in Genetics, 2010 August ; 26(8): 345–352. doi:10.1016/j.tig.2010.05.003.
+
+\bibitem{monsanto}
+  Badran A. H., Guzov V.M., Huai Q., Kemp M. M.,Vishwanath P.,Kain W., Evdokimov A., Moshiri F., Turner K.H.P., Malvar T., Liu D.R.
+  \newblock{{C}ontinuous evolution of Bacillus thuringiensis toxins overcomes insect resistance}
+  \newblock  Nature 533,58-63 (2016) doi:10.1038/nature17938
 \end{thebibliography}

### 2016-07-30 Doc/MSelection/mselection.tex
pois

@@
 \section*{Introduction}
-\chapter{M-Selection}
-\section{Introduction}
 Continuous evolution of bacteriophage is becomming a potent protein engineering
@@
 %IX(minor coat),
-The continuous evolution process produces a viral genome containing a gene
-which has undergone many generations of mutation and selective pressure for
+Continuous evolution rapidly produces a viral genome containing a gene
+which has undergone many generations of mutation and selection for
 a particular property.  The generality of this approach to protein engineering
-is only limited by our ability to create a selection mechanism for the desired
-property and insert an expressable initial form of the gene into the virus.
+is limited by our ability to insert an expressable initial gene into the phage and create a selection mechanism for the desired activity.
 
@@
 \item{}
-A modified viral genome missing a gene required for infectious progeny.
+A modified viral genome replacing a crucial phage gene with a gene to be evolved.
 
 \item{}
-A modified host genome/plasmid to induce an elevated mutation rate.
+A transformed host with inducible mutagenesis.
 
 \item{}
-A modified host genome/plasmid to provide a selection mechanism to provide
-the missing gene product in proportion to the desired activity of the evolving
-gene.
+A host plasmid containing a selection mechanism to provide the crucial phage gene in proportion to the desired activity of the evolving gene.
 \end{itemize}
 
+\section{M-Selection}
 Extremely high in vivo mutation rates are now possible\cite{mutation}.
@@
 
-which simply modulates the rate of mutation by producing a mutagenesis suppressor from the desired activity of the evolving protein.
-Improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.
+The percentage of phage progeny which are exact copies of the parent is given by the Poisson distribution where $$\lambda$$ is the mutation rate per virion (mutation rate/base * 6000bp/ genome).
+The probability of a virion with zero mutations is:
+\vskip 0.2cm
+\begin{eqnarray}
+\label{eq:POC}
+\mathrm{P}\left( 0 \right) = \frac{{e^{ - \lambda } \lambda ^0 }}{{0!}}
+\end{eqnarray}
+
+Taking a phage production average of 100/hour for normal M13 infection requires us to have a clone rate of at least 2\%: Two faithful copies per infection in order to avoid washout.
+A slower flow rate would lower this minimum fraction.
+
+M-selection may allow for, or even require a flow rate which changes over the course of the experiment.  A ramping up of the flow rate would be appropriate
+
+PACE is designed to keep the transit time of host cells through the lagoon to be on the order of one generation.
+However, at very high levels of mutation it seems unlikely that multiple generations of viable E.
+coli mutants can be produced.
+
+If the desired activity of the evolving protein produces a mutagenesis repressor, improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.
 
@@
 
-Effectively, the number of infectious progeny is proportional to the activity
+In PACE, the number of infectious progeny is proportional to the activity
 of the selection mechanism. In principle, if the current mutation of the gene
@@
 
-M-Selection is proposed to the selection mechanism by combining it with
-the mechanism for elevated mutation.
-
 The benefits of M-Selection may include:
 
-The virus is not crippled, but only has an additional gene or domain, similar to
-phage display.
+The virus is not crippled, that is, no critical genes need to be removed from the virion, and no phage genes are required in a host plasmid.
+This avoids the problem associated with the presence of gIII in the host and improves our ability to propagate phage prior to the evolution experiment.
+Of course, care must be taken to ensure that the evolving gene is not discarded before the experiment begins.
+However, M13 with one additional gene(domain) is exactly what has been perfected in phage display.
 
@@
 
-For example, if the selection mechanism stimulated by the desired activity will consist
-of a repressor for the production of mutagenic polymerase II-.
-Positive selection is the result of reduced mutation, resulting in more clones of the
-selected virus relative to its mutants.
-
 As far as we know, modulation of the mutation rate within a particular host has
-not been exploited in this way.  This form of albeit un-natural selection is referred to
-as M-selection, whereby the selection mechanism consists of reducing the rate of mutation
-acting on a particular instance of the viral genome precisely because it is having some
-desired behavior.
+not been exploited in this way.  We refer to this as M-selection, wherein the selection mechanism consists simply of reducing the rate of mutation.
 
@@
 \vskip 0.2cm
-
-\subsection{Difficulties of working with crippled Virus genome}
-
-\begin{eqnarray}
-\label{eq:schemeP}
-	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
-\end{eqnarray}
-
-\subsection*{Multiplicity of Infection}
-
-\vskip 0.2cm
-\begin{eqnarray}
-\label{eq:MOI}
-\mathrm{P}\left( x \right) = \frac{{e^{ - \lambda } \lambda ^x }}{{x!}}
-\end{eqnarray}
 

### 2016-07-31 Doc/MSelection/mselection.tex
merged editing

@@
 \section*{Introduction}
-Continuous evolution of bacteriophage is becomming a potent protein engineering
-tool\cite{pace}\cite{monsanto}.
+Continuous evolution of bacteriophage has the potential to become a potent protein engineering tool\cite{pace}\cite{monsanto}.
 For example, PACE - Phage Assisted Continuous Evolution described by Esvelt,Carlson and Lui. 
@@
 \section{M-Selection}
-Extremely high in vivo mutation rates are now possible\cite{mutation}.
-With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction.
+Extremely high, yet controllable, in-vivo mutation rates are now possible\cite{mutation}.
+This mutagenesis can be generated and controlled within an individual host cell.
+We suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied and the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by proportionally more faithful reproduction.
+We propose calling this mechanism Mutagenesis-selection or M-selection.
+
+his technology maked With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction.
 The modulation of mutation rate within each host cell means that one of these wildly mutating strains could find an environment for faithful reproduction if it should stumble upon a promising mutation in the evolving gene.

### 2016-07-31 Doc/MSelection/mselection.tex
remove unnecessary formatting

@@
 % Template for PLoS
+%M13 genes
+%I(assembly),
+%II(replication init),
+%X (DNA accumulation, inside gII) 
+%III(fusion),
+%IV(assembly),
+%V(destabilizing protein),
+%VI(minor coat),
+%VII(minor coat),
+%VIII(major coat),
+%IX(minor coat),
 %subsubsection is 3rd level heading, PLOS doen't support 4th level.
@@
 \\
-Peter Reintjes\textsuperscript{1}
+Peter Reintjes%\textsuperscript{1,2}
 \\
 \bigskip
-\textbf{1} Museum of Life and Science, Durham, North Carolina, USA
+\textbf{Museum of Life and Science}
+\newline
+\bigskip
+\textbf{Innatrix, Inc.}
+%\textbf{1} Museum of Life and Science, Durham, North Carolina, USA
 \bigskip
@@
 Engineering novel proteins via continuous evolution of bacteriophage currently requires the host bacteria to be transformed with two additional functions: A mutagenesis vector to provide an elevated rate of viral mutation, and a selection mechanism to increase the number of infectious progeny for the individual encoding the improved product\cite{pace}.
-I propose using a mutagenesis suppressor as the selection mechanism whereby the desired activity results in a higher number of clones, thus rewarding the genotype with more faithful progeny.
+I propose using a mutagenesis suppressor as the selection mechanism whereby the desired activity results in a higher number of clones, rewarding the genotype by increasing the number of identical progeny.
+As far as I know, modulation of the mutation rate within a particular host has not been exploited in this way.
 
@@
 
-\linenumbers
+%\linenumbers
 
@@
 For example, PACE - Phage Assisted Continuous Evolution described by Esvelt,Carlson and Lui. 
-%M13 genes
-%I(assembly),
-%II(replication init),
-%X (DNA accumulation, inside gII) 
-%III(fusion),
-%IV(assembly),
-%V(destabilizing protein),
-%VI(minor coat),
-%VII(minor coat),
-%VIII(major coat),
-%IX(minor coat),
 Continuous evolution rapidly produces a viral genome containing a gene
@@
 This mutagenesis can be generated and controlled within an individual host cell.
-We suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied and the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by proportionally more faithful reproduction.
-We propose calling this mechanism Mutagenesis-selection or M-selection.
+I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied and the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by proportionally more faithful reproduction.
+I propose calling this mechanism Mutagenesis-selection or M-selection.
 
-his technology maked With M-selection, these high rates will correspond to little or none of the desired activity, these lines of descent will be subject to catastrophic levels of mutation and low probabilities of faithful reproduction.
-The modulation of mutation rate within each host cell means that one of these wildly mutating strains could find an environment for faithful reproduction if it should stumble upon a promising mutation in the evolving gene.
+High rates of mutation will be associated with little or none of the desired activity, and these lines of descent (both E.
+coli and phage) will be subject to catastrophic levels of mutation and low viability.
 
-The percentage of phage progeny which are exact copies of the parent is given by the Poisson distribution where $$\lambda$$ is the mutation rate per virion (mutation rate/base * 6000bp/ genome).
+The percentage of phage progeny which are exact copies of the parent is given by the Poisson distribution where $\lambda$ is the mutation rate per virion (mutation rate/base * 6kbp/ genome).
 The probability of a virion with zero mutations is:
@@
 
-Taking a phage production average of 100/hour for normal M13 infection requires us to have a clone rate of at least 2\%: Two faithful copies per infection in order to avoid washout.
-A slower flow rate would lower this minimum fraction.
+PACE is designed to keep the transit time of host cells through the lagoon on the order of one generation.
+However, at very high levels of mutation it seems unlikely that succeeding generations of E.
+coli mutants or phage will remain viable.
+Furthermore, the continuous flow environment ensures that there can be no other selection mechanism beyond the number of faithful copies of a particular genome.
+The obvious exception to this would be a colony or biofilm which remains somehow attached in the lagoon, but the population size for such entities would be limited by the surface areas within the lagoon, and not the volume.
 
-M-selection may allow for, or even require a flow rate which changes over the course of the experiment.  A ramping up of the flow rate would be appropriate
+Taking a phage production average of 100/hour for normal M13 infection requires us to have a cloning rate of at least 2\%: Two faithful copies per infection in order to avoid washout.
+The previous consideration of high mutation rates would seem to allow slower flow rates, which will lower this minimum fraction.
 
-PACE is designed to keep the transit time of host cells through the lagoon to be on the order of one generation.
-However, at very high levels of mutation it seems unlikely that multiple generations of viable E.
-coli mutants can be produced.
-
-If the desired activity of the evolving protein produces a mutagenesis repressor, improvements in the desired activity will result in a higher percentage of faithful copies of the parent genome in the succeeding generation.
+An important difference with PACE is that mutagenesis is not induced externally and therefore in uninfected cells, so there may be a minimum number of faithful copies produced before the mutagenesis products are expressed by the infected host.
+It may prove useful to retain the arabinose induced mutagenesis as a way of achieving a baseline mutation rate in the lagoon.
 
@@
 
-In PACE, the number of infectious progeny is proportional to the activity
-of the selection mechanism. In principle, if the current mutation of the gene
-produces a product with none of the desired activity, none of the infectivity
-protein will be produced by the host, and there will be no infectous progeny.
-This process involves a continuous flow of host cells through a reaction vessel
-and so any genotype producing less than two infectious virions per host will
-be washed out.
+A possible drawback to this technique is that the mutation space is less agressively searched as the desired activity improves.
+But there is no particular reason that the modulation of mutation rate necessary to produce the population shift will be a low (e.g.
+natural) mutation rate.
 
-The benefits of M-Selection may include:
+\section{Possible Benefits of M-Selection}
+\begin{enumerate}
+\item{The virus is not crippled, that is, no critical genes need to be removed from the virion which will improve our ability to propagate phage prior to the evolution experiment.
+Of course, as usual, care must be taken to ensure that the evolving gene is not discarded before the experiment begins.
+However, M13 with one additional gene(domain) is exactly what has been perfected in phage display.}
+\item{No phage genes are required in a host plasmid.  This avoids the problem associated with the presence of gIII in the host.}
+\end{enumerate}
 
-The virus is not crippled, that is, no critical genes need to be removed from the virion, and no phage genes are required in a host plasmid.
-This avoids the problem associated with the presence of gIII in the host and improves our ability to propagate phage prior to the evolution experiment.
-Of course, care must be taken to ensure that the evolving gene is not discarded before the experiment begins.
-However, M13 with one additional gene(domain) is exactly what has been perfected in phage display.
+The nominal (high) mutation rate can be chosen to guarantee that there will be a minimum number of non-mutant copies produced for every infection.
+This percentage of clones will then increase as the mutation rate is reduced.
 
-The nominal (high) mutation rate can be chosen to guarantee that there will be
-a minimum number of non-mutant copies produced for every infection.  This percentage of clones will then increase as the mutation rate is reduced.
-
-As far as we know, modulation of the mutation rate within a particular host has
-not been exploited in this way.  We refer to this as M-selection, wherein the selection mechanism consists simply of reducing the rate of mutation.
-
-Figure 1 shows how the combined mutagenic and selection plasmid takes the place of the AP and MP in PACE.
+Figure 1 shows the combined mutagenic/selection mechanism replacing the AP and MP in PACE.
 
@@
 Percentage of activated genes due to Phage Shock promoter.
-Number of polIII-  (mutagenic polymerase) transcripts times error rate due to mutagenic polymerase.
-
+Number of polIII-(mutagenic polymerase) transcripts times error rate due to mutagenic polymerase.
 Error rate of normal polII, polIII
 
-Number of phage particles produced np(0-6 min) = 0 (six minutes before any phage appear),
-then production between 75-200 per hour.
+Number of phage particles produced np(0-6 min) = 0 (six minutes before any phage appears), then production between 75-200 per hour.
 
-Initial single-strand (virion) DNA production occurs in non-mutagenic environment, then mutants appear along with polII- production, until suppression of polII- occurs due to binding.
-Binding ?
-anti-sense RNA transcription ?
-polII-  repression ?
-reduction in mutations.
+Some initial single-strand (virion) DNA production will occur in non-mutagenic environment, then mutants appear, until suppression of mutagenesis due to binding(desired activity).
 
 % Results and Discussion can be combined.
-\section*{Results}
-
-\subsection*{Comparison with Sexual/Asexual Reproduction}
+%\section*{Results}
+\section*{Comparison with Sexual/Asexual Reproduction}
 Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copied of its genome.
 For organisms with access to both forms, the percentage of offspring produced by asexual reproduction would be comparable to our percentage of true copies.
-\subsubsection*{Yeast}
+\subsection*{Yeast}
 Although yeast has access to both sexual and asexual reproduction, the (time) cost of sexual reproduction is high and it is generally invoked under conditions of stress, where an increase in genetic variety of offspring.
-\subsubsection*{Daphnia} 
+\subsection*{Daphnia} 
 In Daphnia, longevity and healthy survival to the end of the season result in so called 'winter eggs' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
@@
 
-\begin{enumerate}
-	\item{react}
-	\item{diffuse free particles}
-	\item{increment time by dt and go to 1}
-\end{enumerate}
-
-\subsection*{Using RNA Antisense.}
+\section{Methods for Lowering the Mutation Rate}
+\subsection*{RNA Antisense.}
 
@@
 
-RNA anti-sense of gIV could be utilized to block the primary phage shock promotor which was the primary trigger activating the high level of mutation in the lagoon.
+RNA anti-sense of gIV could be utilized to block the primary phage shock promotor which is the primary stimulus activating mutagenesis.
 
-Lowering the mutation rate can be achieved in a number of ways.
-
+\subsection*{Other Methods}
 \begin{itemize}
@@
 
-\section*{Discussion}
-Modeling needs work.
-
-\section*{Conclusion}
-
-I think this will work.
-
-\section*{Supporting Information}
+%\section*{Discussion}
+%Modeling needs work.
+%\section*{Conclusion}
+%
+%I think this will work.
+%\section*{Supporting Information}
 
 % Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Fig.}
-\label{S1_Fig}
-{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
+%\paragraph*{S1 Fig.}
+%\label{S1_Fig}
+%{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
 
-\paragraph*{S1 Appendix.}
-\label{S1_Appendix}
-{\bf Algorithm.} The code goes here.
+%\paragraph*{S1 Appendix.}
+%\label{S1_Appendix}
+%{\bf Algorithm.} The code goes here.
 
-\paragraph*{S1 Table.}
-\label{S1_Table}
-{\bf Graph maybe.} The Graph goes here.
+%\paragraph*{S1 Table.}
+%\label{S1_Table}
+%{\bf Graph maybe.} The Graph goes here.
 
-\section*{Acknowledgments}
-So many people to thank.
+%\section*{Acknowledgments}
+%So many people to thank.
 
@@
 Esvelt K. M., Carlson J. C. \& Liu D. R.
-\newblock {A system for the continuous directed evolution of biomolecules}.
+\newblock \textit{A system for the continuous directed evolution of biomolecules}.
 \newblock Nature 472, 499–503 (2011) .
@@
 Badran AH, Liu DR.
-\newblock {{D}evelopment of potent in vivo mutagenesis plasmids with broad mutational spectra}.
+\newblock \textit{{D}evelopment of potent in vivo mutagenesis plasmids with broad mutational spectra}.
 \newblock Nature Communications. 2015;6:8425. doi:10.1038/ncomms9425.
@@
 Lynch M.
-\newblock {{E}volution of the Mutation Rate}.
+\newblock \textit{{E}volution of the Mutation Rate}.
 \newblock Trends in Genetics, 2010 August ; 26(8): 345–352. doi:10.1016/j.tig.2010.05.003.
@@
   Badran A. H., Guzov V.M., Huai Q., Kemp M. M.,Vishwanath P.,Kain W., Evdokimov A., Moshiri F., Turner K.H.P., Malvar T., Liu D.R.
-  \newblock{{C}ontinuous evolution of Bacillus thuringiensis toxins overcomes insect resistance}
+  \newblock \textit{{C}ontinuous evolution of Bacillus thuringiensis toxins overcomes insect resistance}
   \newblock  Nature 533,58-63 (2016) doi:10.1038/nature17938

### 2016-07-31 Doc/MSelection/mselection.tex
review printout

@@
 \setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
 %\rfoot{\thepage/\pageref{lastpage}}
@@
 \fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
+%\lfoot{\sf PLOS}
+\lfoot{\sf NCMLS}
 
@@
 Continuous evolution of bacteriophage has the potential to become a potent protein engineering tool\cite{pace}\cite{monsanto}.
-For example, PACE - Phage Assisted Continuous Evolution described by Esvelt,Carlson and Lui. 
 Continuous evolution rapidly produces a viral genome containing a gene
@@
 \item{}
-A modified viral genome replacing a crucial phage gene with a gene to be evolved.
+A modified viral genome replacing a crucial phage gene with the gene to be evolved.
 
@@
 This mutagenesis can be generated and controlled within an individual host cell.
-I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied and the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by proportionally more faithful reproduction.
+I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied: the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by more faithful reproduction.
 I propose calling this mechanism Mutagenesis-selection or M-selection.
-
-High rates of mutation will be associated with little or none of the desired activity, and these lines of descent (both E.
-coli and phage) will be subject to catastrophic levels of mutation and low viability.
 
@@
 A possible drawback to this technique is that the mutation space is less agressively searched as the desired activity improves.
-But there is no particular reason that the modulation of mutation rate necessary to produce the population shift will be a low (e.g.
+But there is no particular reason that the modulation of mutation rate necessary to produce the population shift will be a particularly low (e.g.
 natural) mutation rate.
@@
 \item{No phage genes are required in a host plasmid.  This avoids the problem associated with the presence of gIII in the host.}
+\item{High rates of mutation will be associated with individuals exhibiting little or none of the desired activity, and these lines of descent (both E.
+coli and phage) will be subject to catastrophic levels of mutation and low viability.
+This may reduce the importance of a carefully controlled flow rate.}
 \end{enumerate}
@@
            }}
-\put(35,0){\boxtwo{MUTAGENESIS\ }{\ \ \ Components}}
+\put(35,0){\boxtwo{MUTAGENIC}{\ \ Components}}
 
-\put(130,15){\vector(1,0){30}}
+\put(134,15){\vector(1,0){26}}
 \put(180,15){\circleone{40}{\Large{\ \ F1}}}
@@
 
-\subsection*{Population Dyamics of Bacteriophage}
+\subsection*{Population Dyamics of Bacteriophage (Under Constuction)}
 \subsubsection*{Clones versus Mutants}
-Although it may sounds like the title of a recent movie, this precisely states the selection mechanism of of mutation.
-
+Although it may sounds like the title of a bad sci-fi movie, this precisely states the selection mechanism of of mutation.
+\newline
 Percentage of activated genes due to Phage Shock promoter.
+\newline
 Number of polIII-(mutagenic polymerase) transcripts times error rate due to mutagenic polymerase.
-Error rate of normal polII, polIII
-
+\newline
+Error rate of normal polII, polIII?
+\newline
 Number of phage particles produced np(0-6 min) = 0 (six minutes before any phage appears), then production between 75-200 per hour.
-
+\newline
 Some initial single-strand (virion) DNA production will occur in non-mutagenic environment, then mutants appear, until suppression of mutagenesis due to binding(desired activity).

### 2016-08-01 Doc/MSelection/mselection.tex
poisson of lamda=4
richards edits

@@
 I propose using a mutagenesis suppressor as the selection mechanism whereby the desired activity results in a higher number of clones, rewarding the genotype by increasing the number of identical progeny.
-As far as I know, modulation of the mutation rate within a particular host has not been exploited in this way.
 
@@
 
-Specifically, the PACE system requires:
+Specifically, the Phage Assisted Continuous Evolution (PACE)\cite{pace} system requires:
 \begin{itemize}
@@
 
-The percentage of phage progeny which are exact copies of the parent is given by the Poisson distribution where $\lambda$ is the mutation rate per virion (mutation rate/base * 6kbp/ genome).
-The probability of a virion with zero mutations is:
+We require 2 clones per generation to prevent washout of a genotype.
+Assuming phage production of 100/hour and a lagoon transit time of one hour, we calculate calculate a mutation rate of about 4 mutations per phage genome.
+The percentage of phage progeny which are exact copies of the parent is given by the Poisson distribution where $\mu = 0$ is the expected number of mutations and $\lambda$ is the mutation rate per virion (mutation rate/base * 6kbp/ genome).
+A mutation rate of $\lambda = 4$ gives a 1.8\% probability of zero mutations.
 \vskip 0.2cm
@@
 \label{eq:POC}
-\mathrm{P}\left( 0 \right) = \frac{{e^{ - \lambda } \lambda ^0 }}{{0!}}
+\mathrm{P}\left( \mu, \lambda \right) = \frac{{e^{ - \lambda } \lambda ^ \mu }}{{\mu!}}
+ \rightarrow P(0,4) = \frac{{e^{ - 4 }}}{{1}} = .018
 \end{eqnarray}
@@
 
-Taking a phage production average of 100/hour for normal M13 infection requires us to have a cloning rate of at least 2\%: Two faithful copies per infection in order to avoid washout.
+Taking a phage production average of 100/hour for normal M13 infection requires us to have a cloning rate of at least 2\%: two faithful copies per infection in order to avoid washout.
 The previous consideration of high mutation rates would seem to allow slower flow rates, which will lower this minimum fraction.
@@
 An important difference with PACE is that mutagenesis is not induced externally and therefore in uninfected cells, so there may be a minimum number of faithful copies produced before the mutagenesis products are expressed by the infected host.
-It may prove useful to retain the arabinose induced mutagenesis as a way of achieving a baseline mutation rate in the lagoon.
+It may prove useful to retain the arabinose-induced mutagenesis as a way of achieving a baseline mutation rate in the lagoon.
 
-As the evolving protein approaches a high level of desired activity, the proportionally lowered mutation rate will ensure that more viral progeny contain exact copies of the parent genotype while a baseline mutation rate ensures that our mechanism to lower the mutation rate will not result in stagnation.
+As the evolving protein approaches a high level of desired activity, the proportionally lowered mutation rate will ensure that more viral progeny contain exact copies of the parent genotype, while a baseline mutation rate ensures that our mechanism to lower the mutation rate will not result in stagnation.
 
@@
 Of course, as usual, care must be taken to ensure that the evolving gene is not discarded before the experiment begins.
-However, M13 with one additional gene(domain) is exactly what has been perfected in phage display.}
+However, M13 with one additional gene or domain is exactly what has been perfected in phage display.}
 \item{No phage genes are required in a host plasmid.  This avoids the problem associated with the presence of gIII in the host.}
@@
 
-Figure 1 shows the combined mutagenic/selection mechanism replacing the AP and MP in PACE.
+Figure 1 shows the combined mutagenic/selection mechanism replacing the Mutagenesis Plasmid(MP) and Accessory Plasmid (AP) in PACE.
 
@@
            }}
-\put(35,0){\boxtwo{MUTAGENIC}{\ \ Components}}
+\put(35,0){\boxtwo{Mutagenic}{\ \ Components}}
 
@@
 
-\subsection*{Population Dyamics of Bacteriophage (Under Constuction)}
+\subsection*{Population Dyamics of Bacteriophage (Under Construction)}
 \subsubsection*{Clones versus Mutants}
@@
 \newline
-Some initial single-strand (virion) DNA production will occur in non-mutagenic environment, then mutants appear, until suppression of mutagenesis due to binding(desired activity).
+Some initial single-strand (virion) DNA production will occur in non-mutagenic environment, then mutants appear, until suppression of mutagenesis due to binding (e.g. the desired activity).
 
@@
 \section*{Comparison with Sexual/Asexual Reproduction}
-Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copied of its genome.
+Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copies of its genome.
 For organisms with access to both forms, the percentage of offspring produced by asexual reproduction would be comparable to our percentage of true copies.
 \subsection*{Yeast}
-Although yeast has access to both sexual and asexual reproduction, the (time) cost of sexual reproduction is high and it is generally invoked under conditions of stress, where an increase in genetic variety of offspring.
+Although yeast has access to both sexual and asexual reproduction, the time and energy cost of sexual reproduction is high and it is generally induced by stress, indicating a need for more genetic diversity.
 \subsection*{Daphnia} 
-In Daphnia, longevity and healthy survival to the end of the season result in so called 'winter eggs' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
+In Daphnia, longevity and healthy survival to the end of the season result in so-called 'winter eggs' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
 The cost differential between sexual and asexual reproduction is low and each generation represents a mix of both.
@@
 
-The components for the broad spectrum, high level mutagenisis described in Badran and Liu\cite{mutation} are variations (often single mutation variants) of the functional domains of high fidelity DNA replication and as such are not suitable targets for RNA anti-sense interference.
+The components for the broad-spectrum, high-level mutagenisis described in Badran and Liu\cite{mutation} are variations (often single mutation variants) of the functional domains of high-fidelity DNA replication, and as such are not suitable targets for RNA anti-sense interference.
 Any such anti-sense interference would be likely to affect their normal counterparts which are needed for faithful reproduction.
 
-RNA anti-sense of gIV could be utilized to block the primary phage shock promotor which is the primary stimulus activating mutagenesis.
+RNA anti-sense of gIV could be utilized to block the phage shock promotor which is the primary stimulus that activates mutagenesis.
 

### 2016-08-02 Doc/MSelection/mselection.tex
Lea's comments

@@
 \section{M-Selection}
-Extremely high, yet controllable, in-vivo mutation rates are now possible\cite{mutation}.
+Extremely high, yet controllable, in vivo mutation rates are now possible\cite{mutation}.
 This mutagenesis can be generated and controlled within an individual host cell.
-I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied: the desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplfied by more faithful reproduction.
+I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied: The desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplified by more faithful reproduction.
 I propose calling this mechanism Mutagenesis-selection or M-selection.
@@
 \begin{enumerate}
-\item{The virus is not crippled, that is, no critical genes need to be removed from the virion which will improve our ability to propagate phage prior to the evolution experiment.
-Of course, as usual, care must be taken to ensure that the evolving gene is not discarded before the experiment begins.
-However, M13 with one additional gene or domain is exactly what has been perfected in phage display.}
+\item{The virus is not crippled, that is, no critical genes are removed from the virion which should improve our ability to propagate phage prior to the evolution experiment.
+Care must be taken to ensure that the evolving gene is not discarded, but M13 plus one additional gene or domain is exactly what has been perfected in phage display.}
 \item{No phage genes are required in a host plasmid.  This avoids the problem associated with the presence of gIII in the host.}
@@
 This may reduce the importance of a carefully controlled flow rate.}
+\item{Clonal Interference\cite{dynamics}, the inabiltiy of a single genotype to gain sufficient advantage to become fixed in a population due to weak sellection pressure is viewed as a problem in experimental evolution (one mutant has an important beneficial mutation, but also has a bad mutation hitch-hiking along which attenuates the advantage as compared to a sibling with neither mutation.}
+
 \end{enumerate}
@@
 
-\put(90,90){\vector(0,-1){15}}
-\put(90,45){\vector(0,-1){14}}
+\put(90,90){\vector(0,-1){12}}
+\put(90,56){\vector(0,-1){26}}
 
@@
 \put(0,15){\circleone{40}{\Large{\ \ P1}}}
-\put(20,15){\vector(1,0){30}}
+\put(0,35){\vector(0,1){16}}
+\put(10,40){\small Protein}
+\put(20,15){\vector(1,0){34}}
+\put(25,18){DNA}
 
 \put(35,90){\boxtwo{\ Phage Shock}{\ \ \ Promoter}}
-\put(-20,55){
+\put(-20,62){
 \fbox{%
@@
 }}
-\put(64,60){\vector(1,0){18}}
-\put(83,57){\fbox{\begin{minipage}{0.05in}
-		\hfill\vspace{0.30in}
+\put(64,70){\vector(1,0){18}}
+\put(83,65){\fbox{\begin{minipage}{0.05in}
+		\hfill\vspace{0.20in}
             \end{minipage}
            }}
-\put(35,0){\boxtwo{Mutagenic}{\ \ Components}}
+\put(35,0){\boxtwo{\ \ Mutagenic}{\ \ Components}}
 
@@
 \section{Comparison with Natural Selection}
-Selection pressure for an optimal mutation rate has its own natural history\cite{evomut} although for settled phyla we see a certain natural mutation rates.
-Bacteriophage and multicellular eukaryote mutation rates are higher than bacteria and single cell eukaryotes (around 0.003).
+Selection pressure for an optimal mutation rate has its own natural history\cite{evomut} although for settled phyla we see consistent\cite{drake91}, low mutation rates.
+Bacteriophage $6.4 X 10^-4$, and multicellular eukaryote mutation rates are higher than bacteria $4 X 10^-11$ and single cell eukaryotes $3 X 10^-10$.
 
@@
 
+\section*{Comparison with Sexual/Asexual Reproduction}
+Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copies of its genome.
+For organisms with access to both forms, the percentage of offspring produced by asexual reproduction would be comparable to our percentage of true copies.
+\subsection*{Yeast}
+Although yeast has access to both sexual and asexual reproduction, the time and energy cost of sexual reproduction is high and it is generally induced by stress, indicating a need for more genetic diversity.
+\subsection*{Daphnia} 
+In Daphnia, longevity and healthy survival to the end of the season result in so-called ``winter eggs'' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
+The cost differential between sexual and asexual reproduction is low and each generation represents a mix of both.
+
 \section*{Materials and Methods}
@@
 \subsubsection*{Clones versus Mutants}
-Although it may sounds like the title of a bad sci-fi movie, this precisely states the selection mechanism of of mutation.
+Although it may sounds like the title of a bad sci-fi movie, this precisely states the selection mechanism of mutation.
 \newline
@@
 %\section*{Results}
-\section*{Comparison with Sexual/Asexual Reproduction}
-Sexual reproduction is not the same as an elevated mutation rate, but it shares the characteristic that even a perfectly fit individual will not be producing faithful copies of its genome.
-For organisms with access to both forms, the percentage of offspring produced by asexual reproduction would be comparable to our percentage of true copies.
-\subsection*{Yeast}
-Although yeast has access to both sexual and asexual reproduction, the time and energy cost of sexual reproduction is high and it is generally induced by stress, indicating a need for more genetic diversity.
-\subsection*{Daphnia} 
-In Daphnia, longevity and healthy survival to the end of the season result in so-called 'winter eggs' which reproduce asexually, providing a percentage of offspring which are clones of the parent.
-The cost differential between sexual and asexual reproduction is low and each generation represents a mix of both.
 
-\section{Methods for Lowering the Mutation Rate}
-\subsection*{RNA Antisense.}
+\subsection{Methods for Lowering the Mutation Rate}
+\subsubsection*{RNA Anti-sense}
 
@@
 
-\subsection*{Other Methods}
+\subsubsection*{Other Methods}
 \begin{itemize}
-	\item Supress pIV production or otherwise block phage shock promotor
-	\item Selectively block downstream components of mutagenesis
+	\item Suppress pIV production or otherwise block phage shock promotor.
+	\item Selectively block downstream components of mutagenesis.
 	\item Enhance production of wild-type high-fidelity replication components.
@@
 \newblock \textit{A system for the continuous directed evolution of biomolecules}.
-\newblock Nature 472, 499–503 (2011) .
+\newblock Nature 472, 499–503 (2011).
 
@@
 
+\bibitem{dynamics}
+Barrick, Jeffrey E., Lenski, Richard E.
+\newblock \textit{{G}enome Dynamics during Experimental Evolution}
+\newblock  Nat Rev Genet, 2013 December, 14(12), pp. 827-839, doi:10.1038/nrg3564
+
 \bibitem{evomut}
@@
 \newblock \textit{{E}volution of the Mutation Rate}.
-\newblock Trends in Genetics, 2010 August ; 26(8): 345–352. doi:10.1016/j.tig.2010.05.003.
+\newblock Trends in Genetics, 2010 August; 26(8): 345–352. doi:10.1016/j.tig.2010.05.003.
 
@@
   Badran A. H., Guzov V.M., Huai Q., Kemp M. M.,Vishwanath P.,Kain W., Evdokimov A., Moshiri F., Turner K.H.P., Malvar T., Liu D.R.
-  \newblock \textit{{C}ontinuous evolution of Bacillus thuringiensis toxins overcomes insect resistance}
-  \newblock  Nature 533,58-63 (2016) doi:10.1038/nature17938
+  \newblock \textit{{C}ontinuous evolution of Bacillus thuringiensis toxins overcomes insect resistance.}
+  \newblock  Nature 533,58-63 (2016) doi:10.1038/nature17938.
+
 \end{thebibliography}
 
+\fbox{\begin{minipage}{4.0in}
+Bacteriophage M13. The genome size is 6407 bp (4). The
+mutational target is 258 bp of an inserted Escherichia coli
+lacZa sequence (5). The spontaneous mutant frequency is 6.4
+x 10-4 \textit{Drake, 1991}
+\end{minipage}}
+
 \end{document}

### 2016-08-07 Doc/MSelection/mselection.tex
polishing msel

@@
 Extremely high, yet controllable, in vivo mutation rates are now possible\cite{mutation}.
-This mutagenesis can be generated and controlled within an individual host cell.
+This broad-spectrum mutagenesis with a (claimed) 320,000:1 dynamic range is generated and can be controlled within an individual host cell.
 I suggest that if this mutagenesis is initiated  by infection, say with the phage shock promotor but then reduced by the desired activity, the two principal components of evolution will be satisfied: The desired activity will be selected for as the number of phage progeny of the individual exhibiting that activity will be amplified by more faithful reproduction.
 I propose calling this mechanism Mutagenesis-selection or M-selection.
+So far, all PACE-derived procedures induce a uniform mutagenesis in the lagoon: M-selection doesn't even require the small-molecule (Ara) induction mechanism.
 
-We require 2 clones per generation to prevent washout of a genotype.
+We require 2 clones per generation to infect hosts and produce phage to prevent washout of a genotype.
 Assuming phage production of 100/hour and a lagoon transit time of one hour, we calculate calculate a mutation rate of about 4 mutations per phage genome.
@@
 \end{eqnarray}
+
+So the maximum per-base mutation rate that a genome could tolerate and still stay in the lagoon is $6400/4 = 6.25 X 10^-4$.
+Liu et al.
+claim 322,000 times basal mutation rate.
+Basal rate is possibly ambiguous here if they are referring to the basal rate of E.
+coli producing E.
+coli  ($4x10-10$) vs.
+E coli producing M13 phage ($7.2 x 10-7$), but in either case their claimed dynamic range covers the required phage mutation crossover rate required to avoid washout.
 
@@
 An important difference with PACE is that mutagenesis is not induced externally and therefore in uninfected cells, so there may be a minimum number of faithful copies produced before the mutagenesis products are expressed by the infected host.
-It may prove useful to retain the arabinose-induced mutagenesis as a way of achieving a baseline mutation rate in the lagoon.
+It may prove useful to retain the arabinose-induced mutagenesis as a way of achieving a baseline mutation rate in the lagoon, or an ante-chamber for the lagoon where induction can begin before exposure to phage.
 
@@
 This may reduce the importance of a carefully controlled flow rate.}
-\item{Clonal Interference\cite{dynamics}, the inabiltiy of a single genotype to gain sufficient advantage to become fixed in a population due to weak sellection pressure is viewed as a problem in experimental evolution (one mutant has an important beneficial mutation, but also has a bad mutation hitch-hiking along which attenuates the advantage as compared to a sibling with neither mutation.}
+\item{Clonal Interference\cite{dynamics}, the inabiltiy of a single genotype to gain sufficient advantage to become fixed in a population due to weak selection pressure is viewed as a problem in experimental evolution (one mutant has an important beneficial mutation, but also has a bad mutation hitch-hiking along which attenuates the advantage as compared to a sibling with neither mutation.
+In our case, maybe not a disadvantage as it keeps variety in the lagoon.
+Multiple genotypes in the lagoon at sample time would be informative.
+Does weak selection correspond to a flatter fitness landscape?
+I'm making an assumption here that M-selection corresponds to what Drake and others refer to as 'weak selection'.}
+
+\item{We get 'genetic drift' for free, or at the cost of defective interference, anyway. To quote Patent \#9023549 (Havard/PACE):
+\quote{\textit{A major problem with traditional directed evolution, whether in vitro or in vivo, is that libraries that do not contain any functional variants will be entirely lost, and the effort wasted--even if functionality lies only a couple of mutations away.
+The present invention overcomes this problem by allowing "switchable" genetic drift.
+This can be achieved by, for example, providing "free" propagation components to all library members, such as by inducible expression from an anhydrotetracycline-regulated titratable promoter, enabling all variants to infect host cells.
+By providing enough "free" propagation components, but less than the optimal level, any functional variants that arise will produce slightly more of the missing propagation components, produce more infectious progeny, and take over the population.
+}}}
+
+\item{Adding a negative selection component may be simpler than in PACE: Just need something to re-up-modulate the mutation rate as a result of off-target binding. A suppressor supressor?}
 
@@
 The nominal (high) mutation rate can be chosen to guarantee that there will be a minimum number of non-mutant copies produced for every infection.
-This percentage of clones will then increase as the mutation rate is reduced.
+This percentage (of clones) will then increase proportunally as the mutation rate is reduced.
 
+One downside to all this is defective interference from all the propagating mutants with none of the desired activity.
+So they are infecting host cells and crowding out the 'nice clones' as we wait for their lineages to undergo Error Catastrophe.
+It would be good to know how many uninfected host are making it through the lagoon.
+If we have an excess of uninfected host, this would reduce defective interference.
+
+\begin{minipage}{\textwidth}
 Figure 1 shows the combined mutagenic/selection mechanism replacing the Mutagenesis Plasmid(MP) and Accessory Plasmid (AP) in PACE.
@@
 \centerline{\bf Fig 1. M-selection Plasmid}
+\end{minipage}
 \vskip 0.2cm
@@
 
+\bibitem{drake91}
+  Drake J. W., 
+  \newblock \textit{{A} constant rate of spontaneous mutation in DNA-based microbes}, 
+  \newblock  Proc Natl Acad Sci U S A. 1991 Aug 15;88(16):7160-4.
 \end{thebibliography}
@@
 mutational target is 258 bp of an inserted Escherichia coli
-lacZa sequence (5). The spontaneous mutant frequency is 6.4
-x 10-4 \textit{Drake, 1991}
+lacZa sequence (5). The spontaneous mutant frequency is $6.4 x 10-4$ per genome or $7.2 x 10-7$ per base pair \textit{Drake, 1991}
 \end{minipage}}

### 2017-12-13 c.pl
updates

@@ -90,7 +90,7 @@ camera_reset :-
     evostat_directory(Dir),
     Cmd = '/usr/bin/uvcdynctrl',
     config_name(Config,_),    % Hostname or evostat argument
-    concat_atom([Config,'.gpfl'], Settings),
+    concat_atom([Config,'i.gpfl'], Settings),
     concat_atom(['--device=',Device],Option),
     Args = ['-L', Settings, Option],
     plog(resettingCamera(Dir, Cmd, Args)),
@@ -328,7 +328,7 @@ initialise(W, Label:[name]) :->
          call(Label,Components),
          findall(_,(component(_,_,Obj),free(Obj)),_), % Clear out previous
 	 maplist(create(@gui), Components),
-         setup_web_values,
+	 setup_web_values,
 	 initPID,                        % Start PID controllers
          send(@action?members, for_all,
 	      if(@arg1?value==pIDon,message(@arg1, active, @off))),

### 2017-12-13 et.pl

@@ -45,13 +45,16 @@ term_expansion(iface(Type,PType,Vars), []) :-
        ( pull(S2,N2:name) :-> "Pull value from Arduino"::
 	                      send(S2,converse,N2),
 			      get(S2,reply, Reply),
-	                      parse_reply_arg1(Reply, N2, V2),
-			      nonvar(N2),
-			      nonvar(V2),
-	                      send(S2, N2, V2),
-                              !,
-	                      S2 = @Npull, flog(pull(Npull,N2,V2,succeeded))
-			      ; S2 = @Npull,flog(pull(Npull,N2,failed))
+	                      ( parse_reply_arg1(Reply, N2, V2),
+			        nonvar(N2),
+			        nonvar(V2),
+	                        send(S2, N2, V2)
+			      -> S2 = @Npull,
+				 flog(pull(Npull,N2,V2,succeeded)),
+				 retractall(webValue(S2,N2,_)),
+				 assert(webValue(S2,N2,V2))
+			       ; S2 = @Npull,flog(pull(Npull,N2,failed))
+			      )
        ),
        ( update(US) :->
 	 "Get r/o and push r/w values to Device"::

### 2017-12-13 prephtml.pl

@@ -38,8 +38,9 @@ floatfmt(N,[A|T],T) :-
 % Internal temperatures integral tenths of degrees C.
 
 float_tenths(Obj,Thing) --> [Display],
-    { get(Obj,Thing,InTenths),
-
+      {
+      get(Obj,Thing,InTenths),
+      (Thing = tt -> plog(get(Obj,Thing,InTenths)) ; true),
       refresh(Obj,Thing,InTenths),
       DispTemp is float(InTenths)/10.0,
       format(atom(Display), '~4G', [DispTemp]) }.

### 2017-12-14 Doc/Cameras/alevel.tex
alevel documentation

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{changepage}
\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{cite}
\usepackage{nameref}
\usepackage[pdftex,
            pdfauthor={Peter Reintjes},
            pdftitle={EvoStat Level Monitoring System},
            pdfsubject={Level Monitoring},
            pdfkeywords={OpenCV Python},
            pdfproducer={Latex with hyperref},
            pdfcreator={pdflatex}]{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage[right]{lineno}
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
\usepackage[table]{xcolor}
\usepackage{array}
\newcolumntype{+}{!{\vrule width 2pt}}
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

\usepackage{enumitem}
\setlist{nosep}
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}
% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx,wrapfig}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
%\rfoot{\thepage/\pageref{lastpage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
% customize
\lfoot{\sf Innatrix Internal Document}

\begin{document}

\title{EvoStat Level Monitoring System}
\author{Peter Reintjes}
\date{\today}
\maketitle

\vspace*{0.2in}
\begin{flushleft}
\includegraphics[scale=0.5]{{phagestat}.jpg}
\end{flushleft}

{\it The EvoStat liquid-level control system consists of two distict parts: Monitoring and Control, which are accomplished by separate programs.  This report focuses on the monitoring program (alevel.py), a Python/OpenCV\cite{opencv} program which observes and reports the levels of vessels within the main EvoStat chamber.  The main EvoStat control program requires this level information to maintain volumes, cell density, and flow rates in the various chambers to support experiments in the continuous evolution of bacteriophage\cite{pace}.}

\section*{Introduction}

The {\bf alevel.py} program is a Python program which takes images from a digital camera and determines the individual liquid levels in transparent containers within the image.  The EvoStat system uses the program to report the liquid levels of up to five vessels, one 2-liter host-cell cultivator and four 250mL lagoons.

\section*{Level Detection in the EvoStat}
Whenever a level reading is to be taken, {bf alevel.py} reads a text file {\bf .previous}) which specifies for each vessel:

\begin{enumerate}[itemsep=1pt, topsep=2pt, partopsep=0pt]
\item Area of the image in which the meniscus appears
\item Recommended intensity amplification (images to sum)
\item Recommended contrast (gain and threshold) settings
\end{enumerate}

\subsection*{Camera Settings}
The unusual (low) light conditions inside the EvoStat may require a camera to be initialized with specific exposure, contrast, and saturation settings. While many cameras have automatic exposure modes, detection can often be improved with custom settings to match the dark conditions and our use of primary colors: red for the reticules, green for the liquid menisci.

One of the legacies of Windows\textregistered and other closed systems is that each camera manufacturer has to produce a program for the (closed) operating system for each of their cameras.  Linux systems have kept up pretty well with a tool called guvcview which knows about many cameras -- especially the cheap ones.  This interactive program can be used to modify settings on many cameras and save those settings in a file with the {\bf .gpfl} suffix.  This file can be used to initialize the camera when the program starts by calling the program {\bf uvcdynctrl}.

After making the necessary image adjustments to the camera at {\bf /dev/video0} with {\bf guvcview} and saving the settings in the file {\bf evostat.gpfl}, the level monitor program (or any other process) can run the following command to restore the camera to these settings:
\begin{verbatim}
       /usr/bin/uvcdynctrl -L evostat.gpfl --device=/dev/video0
\end{verbatim}

\subsection*{Alignment}
It is not necessary to align the camera precisely as long as it has a clear view of the reticule column and that the regions specified for each vessel in the {\bf .previous} file surround a sufficient horizontal region of the meniscus.

\subsubsection*{The Reticules}
The {\bf reticules} are movable red LED assemblies on a black column inside the EvoStat. The upper assembly should be aligned with the target liquid level in the upper, or host-cell cultivation chamber, while the lower assembly should be aligned with the target level for the lagoons.  The EvoStat is designed so that all of the lagoons are on the same level, hence we only need one reticule to provide a reference for all of the lagoon levels.

\begin{wrapfigure}{L}{0.25\textwidth}
\centering
\includegraphics[width=0.23\textwidth,scale=0.2]{reticules.png}
\caption{\label{fig:reticules}Reticules}
\end{wrapfigure}

The placement of the reticule does not limit the level control to this level. Level control is relative to the reticules and may be set to maintain levels at an offset above or below the level of the reticules. In the simplist case, it is convenient to position the reticule assemblies so that the two red bars are placed to surround the desired liquid level.  This would correspond to an offset of zero for the liquid level to be maintained.

\subsubsection*{Vessels}
1) the region of the image where the meniscus will be visible, for each vessel and intensity and contrast settings that have been selected by the user.

\subsection*{Lighting}
Valid levels cannot be obtained if the EvoStat chamber is opened, and therefore exposed to normal room lighting.  While the ability to darken the room can substitute for a closed chamber, we haven't had that luxury since being evicted from the Tod Vision's gardening tool shed in the Genome Sciences Building.

Therefore, whenever the level monitoring program is called by the EvoStat and good level readings are not possible, there is no output produced and the EvoStat control program makes operates as if the readings were unchanged.

\subsection*{Growth Conditions}

\bibliography{alevel}{}
\bibliographystyle{unsrt}
\end{document}

### 2017-12-14 alevel.py
sync alevel

@@ -540,6 +540,8 @@ def imageOut():
     (he,wi,de) = referenceImage.shape
     cv2.putText(referenceImage,time.asctime(time.localtime()),(wi/10,he/2),
                 cv2.FONT_HERSHEY_SIMPLEX, 1, (255,210,200),2)
+    cv2.putText(referenceImage,lasttemp(),(wi/10,4*he/7),
+                cv2.FONT_HERSHEY_SIMPLEX, 0.8, (250,210,255),2)
     cv2.imwrite(imageName,cv2.resize(referenceImage,params['imageSize']))
     movie_file(imageName)
 
@@ -690,6 +692,13 @@ def get_camera() :
 #        time.sleep(1)
     camSettle(3)   # save_frames('sample',10) to create sample set
 
+def lasttemp() :
+    cmd = ['/bin/bash', './lasttemps.sh']
+    p = subprocess.Popen(cmd, stdout=subprocess.PIPE)
+    (result, err) = p.communicate('')
+    result.replace('88.8 C','No Data')
+    return result
+    
 if __name__ == "__main__" :
     global referenceImage
     global positions

### 2017-12-15 Doc/Cameras/alevel.tex
minor edits

@@
 \usepackage{array}
+\usepackage{listings}
 \newcolumntype{+}{!{\vrule width 2pt}}
@@
 
-The {\bf alevel.py} program is a Python program which takes images from a digital camera and determines the individual liquid levels in transparent containers within the image.
+The {\bf alevel.py} program is a Python program which takes an image from a digital camera and determines the individual liquid levels in transparent containers within the image.
 The EvoStat system uses the program to report the liquid levels of up to five vessels, one 2-liter host-cell cultivator and four 250mL lagoons.
@@
 \section*{Level Detection in the EvoStat}
-Whenever a level reading is to be taken, {bf alevel.py} reads a text file {\bf .previous}) which specifies for each vessel:
+Whenever a level reading is to be taken, {\bf alevel.py} reads a text file {\bf .previous} which specifies for each vessel:
 
@@
 \end{enumerate}
+Here is an example of the {\bf .previous} file containing the recommended settings and vessel locations.
+The {\bf reticule} entry refers to the red reference LEDs used for level calibration.
+The absence of a bounding box indicates that the reticules can be anywhere within the image area.
+\begin{verbatim}
+# (Vessel,oldPt,BBox,Contrast-iter, Contrast-mul, Contrast-off,')
+#               Erode-iter, Dilate-iter, Amplify-iter, Amp-factor')
+# (name, (x,y), ((uy,lx),(ly,rx)), 1, 1.2, -80, 2, 2, 3, 0.7)')
+# lagoon   1, 1.5972, -90, 2, 2, 3, 0.7)]
+[('reticule', 0, 1.4, -80, 180, 1, 4, 3, 0.4),
+ ('host0', (47, 22), ((40, 120), (214, 380)), 1, 1.2, -80, 2, 2, 2, 0.4),
+ ('lagoon1', (84, 71), ((340, 50), (460, 260)), 1, 1.1, -110, 2, 2, 3, 0.7)]
+\end{verbatim}
+
+The file is named {\bf .previous} because the program will try these recommended values for contrast and intensity settings but vary them until it can get a good level reading.
+If the user specified the {\bf update} option when invoking {\bf alevel.py} a new file will be written with the contrast and intensity values actually used.
+The next time the program is run, these settings may be able to get a good level reading without the need to vary them.
+The program may also adjust the bounding boxes to be smaller and centered on the meniscus that was actually measured.
 
@@
 
-One of the legacies of Windows\textregistered and other closed systems is that each camera manufacturer has to produce a program for the (closed) operating system for each of their cameras.
+One of the legacies of
+Windows\textsuperscript{\textregistered}
+and other closed systems is that each camera manufacturer has to produce a program for the (closed) operating system for each of their cameras.
 Linux systems have kept up pretty well with a tool called guvcview which knows about many cameras -- especially the cheap ones.
@@
 
+Although the EvoStat programs are otherwise portable between Linux and
+Windows\textsuperscript{\textregistered}
+systems, we know of no comparable program that would make this method
+of camera initialization available on
+Windows\textsuperscript{\textregistered}
+systems.
+
 \subsection*{Alignment}
@@
 The upper assembly should be aligned with the target liquid level in the upper, or host-cell cultivation chamber, while the lower assembly should be aligned with the target level for the lagoons.
-The EvoStat is designed so that all of the lagoons are on the same level, hence we only need one reticule to provide a reference for all of the lagoon levels.
+The EvoStat is designed so that all of the lagoons are on the same level, hence we only need one reticule to provide a reference for the lagoon levels.
 
@@
 
-The placement of the reticule does not limit the level control to this level.
-Level control is relative to the reticules and may be set to maintain levels at an offset above or below the level of the reticules.
-In the simplist case, it is convenient to position the reticule assemblies so that the two red bars are placed to surround the desired liquid level.
-This would correspond to an offset of zero for the liquid level to be maintained.
+Reticule placement does not directly specify the level being controlled.
+Level control is relative to the reticules and may be set to maintain a level at an offset above or below the level of the reticule.
+In the simplist case, however, it is convenient to position the reticule assembly so that the two red bars surround (above and below) the desired liquid level.
+This corresponds to a zero offset for the liquid level to be maintained.
 
 \subsubsection*{Vessels}
-1) the region of the image where the meniscus will be visible, for each vessel and intensity and contrast settings that have been selected by the user.
+
+For each vessel, the
+{\bf .previous}
+file has a bounding box defining the region of the image where the meniscus will be visible, and intensity and contrast settings. Once
+{\bf alevel.py}
+reads this file, it will try to detect the horizontal liquid level in each of these regions using the contrast and intensity settings found in the file.
+However, if this fails it will vary each of these parameters by positive and negative factors in 10\% increments until it gets a good reading (or a specified number of attempts have been made).
+With the {\bf update} option,
+{\bf alevel.py}
+will re-write the new values to the
+{\bf .previous}
+file.
 
 \subsection*{Lighting}
-Valid levels cannot be obtained if the EvoStat chamber is opened, and therefore exposed to normal room lighting.
+
+The various EvoStat vessel holders shine light from bright green LEDs up into the liquid containers.
+When liquids are clear (e.g.
+uninoculated), this results in a reflection off the meniscus which can be used to determine the liquid level.
+However, the system works most reliably with the turbidity of a growing culture which is the primary mode of EvoStat operation.
+Until the image processing has been improved, we should consider that accurate level readings will only occur when there is active culture in the vessels (e.g.
+not during setup and sterility testing).
+
+Level readings also fail when the EvoStat chamber is opened, and therefore exposed to normal room lighting.
 While the ability to darken the room can substitute for a closed chamber, we haven't had that luxury since being evicted from the Tod Vision's gardening tool shed in the Genome Sciences Building.
 
-Therefore, whenever the level monitoring program is called by the EvoStat and good level readings are not possible, there is no output produced and the EvoStat control program makes operates as if the readings were unchanged.
+If a good level reading is not possible, calls to the level monitoring program return no results and the EvoStat reacts as if the last good reading was repeated.
+As with all other parts of the EvoStat, the system is designed to operate with the current settings through any loss of telemetry.
 
 \subsection*{Growth Conditions}
+Under actual growth conditions, the liquids are turbid and create a lot of scattered light.
+The LEDs below the vessel must be bright enough to completely illuminate the contents.
+This is particularly important for the large, upper chamber because the light source is at the bottom of the 2-liter container.
+In particular, we do not see a high-contrast meniscus when the volume is near the overflow drain (1.5 liter level).
 
+\section*{Next Steps}
+
+Now that the level reporting program is relatively stable, a complete code-review re-write is in order.
+Duplicated code can be eliminated in the horizontal line detection and image amplification subroutines.
+Red/reticule and Green/Level readings had minor differences which were best debugged using separate implementations.
+
+Produce level information in resolution independent coordinates so that they are the same whether a 640x480 or 1280x1024 image is available.
+Or, alternately, provide image resolution information with the data and allow the recieving program to do the coordinate mapping.
+
+\begin{enumerate}[itemsep=1pt, topsep=2pt, partopsep=0pt]
+\item Review/re-write code: 800 lines of Python, can probably be halved.
+\item Make level data resolution independent
+\item Automate camera settings under Windows\textsuperscript{\textregistered}
+\end{enumerate}
+
+\bibliographystyle{unsrt}
 \bibliography{alevel}{}
-\bibliographystyle{unsrt}
+
+\pagebreak
+
+\section{Appendix: alevel.py source code}
+\newcommand*{\SrcPath}{../..}
+\lstinputlisting[language=Python]{\SrcPath/alevel.py}
 \end{document}
+
+\input{\MyPath/chap1/ch1.tex}

### 2020-03-05 Doc/MSelection/mselection.tex
images included
sensitive data purge

@@
 \date{}
-
 % Header and Footer with logo
@@
 \bigskip
-\textbf{Museum of Life and Science}
-\newline
-\bigskip
 \textbf{Innatrix, Inc.}
-%\textbf{1} Museum of Life and Science, Durham, North Carolina, USA
 \bigskip
@@
 
-One downside to all this is defective interference from all the propagating mutants with none of the desired activity.
+The downside to this is defective interference from all the propagating mutants with none of the desired activity.
 So they are infecting host cells and crowding out the 'nice clones' as we wait for their lineages to undergo Error Catastrophe.
@@
 If we have an excess of uninfected host, this would reduce defective interference.
+
+In the first generation, even a fatal mutation will be packaged in a viable virion, so that it will infect a host cell.
+But it will go no further if the fatal mutation results in non-functional coat proteins, origin of replication, packaging signal, or whatever.
+So in the worst case, a fatal mutation results in the production of 100 dead-end virions.
+
+The maximum mutation rate will be experienced by phage that have none of the desired activity and we want this rate to ensure that the accumulation of 5 non-beneficial mutations will result in a fatality.
+The growth of a defectively interfering population is thus limited to 100^5, 10^6.
+This number is reached only if for phage that has accumulated four non-beneficial, non-fatal mutations.
+Presumably these are silent mutations as the phage genome is minimal and highly conserved.
+We should also recognize that a reversion to wild-type phage (shedding the payload gene), will also subject the phage to maxiumum mutation in every generation.
+
+Beneficial mutations resulting in detection of the desired activity will reduce the mutation rate, resulting in more cloned phage, which will inherent this ability to supress the mutation rate.
+
+Even if it is possible to have a series of mutations which result in a viable phage after ten generations, this genotype will not.  For the high-mutation
+all of the different mutations that occur in these 10 generations would have to be non fatal. resulting in 10^10 viable phage mutations.
+
+For a given maximal mutation rate, what are the
+Can we estimate a probability for M13s resistance to mutation.
+Since different types of organisms, and even individual species, have some control over their exposure to mutations.
+
+Perfect replication okay for Artificial Selection
+
+There is at least one aspect of natural evolution which we do not have to replicate in directed evolution.
+The principals of evolution make it clear that perfect DNA replication is a recipe for extinction in any environment that is subject to change.
+But we are clearly going to engineer our continuous evolution environment to be as stable as possible, so future experiments can take advantage of DNA replication and repair mechanisms to allow for perfect replication at the upper end of the usable range.
+
+Phage has a higher natural mutation rate than all types of cells, but our artificial environment does not have to respect
+There is no reason not to utilize the entire range from Mutation Storm to Perfect Replication.  In fact, modifications which can achieve perfect replication.
+
+So we may do something nature never did, which is bring all the tools of replicative fidelity to phage reproduction.
+
+What fraction of the M13 gene can carry non-fatal mutations. Including all the silent mutations, plus those which represent similar amino acid substitutions,
 
@@
 % Results and Discussion can be combined.
-%\section*{Results}
+\section*{Results}
+\includegraphics{{0.0cAMP_M13_MP6-7View}.jpg}
+\newline
+M13 with 0.0 cAMP + MP6 $10^-7$
+\includegraphics{{5mMcAMP_M13_MP6-5View}.jpg}
+\newline
+M13 with 5 milliMolar cAMP + MP6 $10^-5$
+\includegraphics{{50mMcAMP_M13_MP6-5View}.jpg}
+\newline
+M13 with 50 milliMolar cAMP + MP6 $10^-5$
 
@@
   \newblock  Proc Natl Acad Sci U S A. 1991 Aug 15;88(16):7160-4.
+
+\bibitem{domingo2009}
+  Domingo-Calap P, Cuevas JM, Sanjua´n R (2009)
+\newblock The Fitness Effects of Random Mutations in Single-Stranded DNA and RNA Bacteriophages.
+\newblock PLoS Genet 5(11): e1000742. doi:10.1371/journal.pgen.1000742
+
 \end{thebibliography}

### 2020-03-05 alevel.py

@@ -1,5 +1,9 @@
 #!/usr/bin/python -u
 #!C:/cygwin/Python27/python -u
+
+# minlen for green horizontal lines changed from 6 to 10 to avoid reflection
+# minlen for reticules is still 8
+
 from __future__ import print_function
 import sys, os, time, socket, glob, subprocess
 # To get rid of spurious messages from openCV library
@@ -7,7 +11,6 @@ from suppress_stdout_stderr import suppress_stdout_stderr
 import numpy as np
 import cv2
 import cv2.cv as cv
-
 from shutil import copyfile
 
 debug = False
@@ -59,19 +62,19 @@ def plog(str) :
         print("      --"+str, file=sys.stderr)
 
 def termIntList(f,l) :
-    if (l == None ):
+    if (l is None ):
         plog("Empty argument list: " + f)
         return f+"."
     return f+"("+", ".join([str(i) for i in l])+")."
 
 def termPairList(f,l) :
-    if (l == None ):
+    if (l is None ):
         plog("Empty argument list: " + f)
         return f+"."
     return f+"("+", ".join([str(i[0])+":"+str(i[1]) for i in l])+")."
 
 def termTripleList(f,l) :
-    if (l == None ):
+    if (l is None ):
         plog("Empty argument list: " + f)
         return f+"."
     return f+"("+", ".join([str(i[0])+":"+str(i[1])+":"+str(i[2]) for i in l])+")."
@@ -84,7 +87,7 @@ def termNumberList(f,l) :
 
 
 def icheck(image, who) :
-    if (image == None) :
+    if (image is None) :
         plog("Image equal to None in " + who)
         exit(13)
     if (len(image.shape) == 2) :
@@ -117,7 +120,7 @@ def get_previous() :
         plist = eval(open(prevfile, 'r').read())
     except(IOError):
         print("No previous level information")
-    if (plist == None) :
+    if (plist is None) :
         print("creating PLIST")
         (it, sc1, of1, ei, di, ai1, af) = ( 1, 1.85, -80, 2, 2, 4, 0.7 )
         (sc2, of2, ai2) = (1.35, -80, 3)
@@ -146,6 +149,7 @@ def showdb(img, delay=500) :
             exit(0)
 
 def exportImage(image) :
+    global imageName
     if (image != None ) :
         (cy1,cx1,cy2,cx2) = params['cellstatRegion']
[…]

### 2020-03-05 aristotle.pl

@@ -64,12 +64,12 @@ config( [
 %PREVbt_device( lagoon1,  '98:D3:31:20:23:36'). % HC-06 name: cellstat0
 
 % Ari's devices
-bt_device( host0,  '98:D3:31:FB:7A:BF'). % HC-06 name: <new>
-bt_device( lagoon2,  '98:D3:31:FB:A5:15'). % HC-06 name: <new>
+%bt_device( host0,  '98:D3:31:FB:7A:BF'). % HC-06 name: <new>
+%bt_device( lagoon2,  '98:D3:31:FB:A5:15'). % HC-06 name: <new>
 
 % Huxley's devices
-%bt_device( host0,  '98:D3:31:20:23:36'). % HC-06 name: cellstat0
-%bt_device( lagoon2,  '98:D3:31:40:31:C3'). % HC-06 name: lagoon2
+bt_device( host0,  '98:D3:31:20:23:36'). % HC-06 name: cellstat0
+bt_device( lagoon2,  '98:D3:31:40:31:C3'). % HC-06 name: lagoon2
 
 % NEW HOST
 

### 2020-03-05 c.pl

@@ -413,7 +413,7 @@ initialise(W, Label:[name]) :->
          send(W,started),
          send_super(W, open, Location),
 	 plog(normal(launch)),
-	 launch(level),
+	 launch(level), % calls Python program level.py
 	 findall(Person,sendIP(Person),People),
 	 plog(sentIP(People)),
 	 plog(finished(evostat)).
@@ -889,7 +889,8 @@ main(Argv) :-
 %   xrestart and linuxrestart are handled in by crontab (system_actions.sh)
 %
 
-levelrestart :- launch(level), % Should clean up any old instance of process
+% Should clean up any old instance of process
+levelrestart :- launch(level), % runs Python program level.py
 		plog(relaunched(level)).
 
 % findall, maplist vs. repeat,fail,true

### 2020-03-05 clevel.py

@@ -239,55 +239,62 @@ def level_data(img, minlen=12) :
     if (alllines == None) :
         return (0,0)
     maxvert = 0
-    maxvertpos = 0
     numbervert = 0
     centerx = -1
-    state = 0
+    print("VRESULT")
     for lines in alllines:
+        print("  GROUP "+str(len(lines)))
         for l in lines :
             if (l[0] == l[2]) : # Vertical line
-                numbervert = numbervert + 1
                 vlen = l[1]-l[3]
+                numbervert = numbervert + 1
+                print("    Vert at "+str(l[0]) +" len "+str(vlen))
                 if (vlen > maxvert) :
                     centerx = l[0]
                     maxvert = vlen
-                    maxvertpos = l[0]
     report = report + "num vert = "+str(numbervert)+" centerx="+str(centerx)+" max vert="+str(maxvert)
     numberhoriz = 0
     maxy = -1
     miny = 1000
+    leftofcenter = None
+    rightofcenter = None
     if (numbervert == 0) : # Check for one horizontal line
+        print("H1RESULT")
         for lines in alllines:
+            print("  GROUP "+str(len(lines)))
             for l in lines : # Find max Y line (not on the edge)
                 if (l[1] == l[3]) : # Horizontal
                     hcen = (l[0]+l[2])/2
+                    hlen = l[2] - l[0]
                     numberhoriz = numberhoriz + 1
-                if ( l[1]>maxy ) :
-                    maxy = l[1]
-                    leftofcenter = ( hcen, l[1] )
-                    state = 1
-                    report = report + " hl"+str(hcen)
-        return (state, 0)
-    leftofcenter = None
-    rightofcenter = None
-    for lines in alllines:
-        for l in lines : # Find min Y line (not on the edge)
-            if (l[1] == l[3]) : # Horizontal, not near top or bottom
-                numberhoriz = numberhoriz + 1
-                if ( l[1] < h-3 and l[1] > 3 and l[1] < h-3) :
-                    hcen = (l[0]+l[2])/2
-                    if ( (hcen < centerx) and l[1]>maxy ) :
+                    print("    Horz at "+str(hcen)+ " len=",str(hlen))
+                    if ( l[1]>maxy ) :
                         maxy = l[1]
                         leftofcenter = ( hcen, l[1] )
-                        report = report + "left " + str(hcen)
-                    if ( (hcen > centerx) and l[1]<miny ) :
-                        miny = l[1]
-                        rightofcenter = ( hcen, l[1] )
[…]

### 2020-03-05 contrast.py

@@ -1,127 +1,0 @@
-#!/usr/bin/python -u
-#!C:/cygwin/Python27/python.exe -u
-#!C:/Python27/python -u
-
-from __future__ import print_function
-def plog(str) :
-    print(str, file=sys.stderr)
-
-from suppress_stdout_stderr import suppress_stdout_stderr
-import time, sys
-import numpy as np
-import cv2
-import cv2.cv as cv
-from Tkinter import *
-
-gc_color = 1
-gc_iter = None
-gc_scale = None
-gc_offset = None
-gc_camera = None
-
-def show(img) :
-    cv2.imshow("camera", img)
-    if cv.WaitKey(2000) == 27:
-        exit(0)
-    
-def contrast(image, iter=1, scale=2.0, offset=-80) :
-    if (image == None) :
-        plog("contrast called with null Image")
-    for i in range(iter) :
-        plog("Try contrast "+str((iter,scale,offset)))
-        if (image == None) :
-            plog("contrast loop: Image is None")
-        else :
-            show(image)
-        image = cv2.add(cv2.multiply(image,scale),offset)
-        if (image == None) :
-            plog("image(None) after add/mulitply in contrast!")
-        show(image)
-        
-        (ret,img) = cv2.threshold(image, 127, 255, cv2.THRESH_BINARY)
-        if (ret == False) :
-            plog("Thresholding failed?")
-            return None
-    if (img == None) :
-        plog( "img is None after binary threshold in contrast")
-    return(img)
-
-def setred() :
-    global gc_color
-    gc_color = 2
-
-def setblue() :
-    global gc_color
-    gc_color = 0
-
-def setgreen() :
-    global gc_color
-    gc_color = 1
-
-def showContrast() :
-    """Check contrast algorithm using three values from TkInter console"""
-    global gc_color
-    global gc_scale
-    global gc_iter
-    global gc_offset
-    (rval, frame) = gc_camera.read()
-    show(frame)
-    greyimage = frame[:,:,gc_color]
-    s = float(gc_scale.get())/100.0
-    o = gc_offset.get()
-    i = gc_iter.get()
-    show(greyimage)
-    greyimage = contrast(greyimage,iter=i,scale=s,offset=o)
-    show(greyimage)
-
-def makeTkSliders(contrast) :
-    global gc_iter
-    global gc_scale
-    global gc_offset
-    global gc_camera
-    master = Tk()
[…]

### 2020-03-05 cscam.py

@@ -1,681 +1,0 @@
-#!/usr/bin/python -u
-#!C:/cygwin/Python27/python -u
-#!C:/Python27/python -u
-
-from Tkinter import *
-
-gc_iter = None
-gc_scale = None
-gc_offset = None
-ipcam = None
-
-def showContrast() :
-    global ipcam
-    ipcam.checkContrast()
-    
-def makeTkSliders() :
-    global gc_iter
-    global gc_scale
-    global gc_offset
-    master = Tk()
-    master.minsize(250,150)
-    master.title("Contrast Setting")
-    iter_label = Label(master,text="Number of Iterations")
-    iter_label.pack()
-    gc_iter = Scale(master, from_=1, to=4, orient=HORIZONTAL)
-    gc_iter.pack()
-    
-    scale_label = Label(master,text="Scaling Multiplier")
-    scale_label.pack()
-    gc_scale = Scale(master, from_=0, to=200, orient=HORIZONTAL)
-    gc_scale.pack()
-    
-    offset_label = Label(master,text="Negative Offset after Scale")
-    offset_label.pack()
-    gc_offset = Scale(master, from_=-100, to=0, orient=HORIZONTAL)
-    gc_offset.pack()
-    
-    Button(master, text='Show', command=showContrast).pack()
-    Button(master, text='Quit', command=exit).pack()
-    master.mainloop()
-
-import sys, os, time, socket, subprocess, re, traceback
-
-from os  import popen
-import glob
-import base64, urllib2
-from suppress_stdout_stderr import suppress_stdout_stderr
-import numpy as np
-import cv2
-import cv2.cv as cv
-import evocv
-
-def levelTerm(Levels, What) :
-    return What + "( "+", ".join([str(Levels[k]) for k in Levels.keys()])+")."
-
-def on_exit(msg,nLagoons) :
-    print("levels(80, 60, 40, 20).")
-    print "message('" + msg + "')."
-    exit(0)
-
-rbox =  { 'x1':100,'y1':100,'x2':110,'y2':110 }
-bbfp = None
-threshold = 1100
-
-def on_mouse(event, x, y, flags, params):
-    global rbox
-    if event == cv.CV_EVENT_LBUTTONDOWN:
-        rbox['x1'] = x
-        rbox['y1'] = y
-    elif event == cv.CV_EVENT_LBUTTONUP:
-        rbox['x2'] = x
-        rbox['y2'] = y
[…]

### 2020-03-05 greenlines.py

@@ -564,6 +564,10 @@ def processRegion2(image, bbox, function, params, n) :
 if __name__ == "__main__" :
     global referenceImage
     global positions
+    with suppress_stdout_stderr() :
+        os.system("sudo ./usbreset /dev/bus/usb/005/011")
+        time.sleep(1)
+        os.system("/usr/bin/uvcdynctrl -L aristotle.gpfl --device=/dev/video0")
     pfile = open("vary.txt",'w')
     for v in vary(defaultParams) :
         print(v,file=pfile)
@@ -597,7 +601,7 @@ if __name__ == "__main__" :
     rlevels = getLevels(green, 127, (1, 1.28, -80), 2, reticules)
     printRLevels(rlevels)
     imageOut()
-    release()
+#    release()
 #    print(str(vlevels))
 #    print(str(get_positions(vlevels)))
 #    print(str(xydeltay(reticules,get_positions(vlevels))))

### 2020-03-05 haldane.pl

@@ -1,72 +1,0 @@
-config( [
-	 updateCycle(20),
-	 debugpause(100),
-	 numLagoons(2),
-	 cellstatRegion(30,100,200,350),
-	 cellstatScale(1500),
-	 cellstatOffset(100),
-%         cellstatContrast(2,1.4,-80),
-         cellstatContrast(3,2,-80),
-         lagoonRegion(264,45,468,637),
-	 lagoonHeight(120),    % divisor for levelScale
-	 lagoonWidth(60),
-%         lagoonContrast(2,1.35,-70),
-         lagoonContrast(2,1.76,-80),
-	 levelScale(140),
-	 levelOffset(10),
-	 levelScale(100),   % 100 gives level as percentage of lagoonHeight
-	 darkness(60),
-	 camera(indoor),
-	 mac(0),
-	 frames(120),
-	 rotate(90),
-	 defaultIP('192.168.2.139'),  % belongs in snapshot
-         userpwd('&user=scrapsec&pwd=lakewould'),
-         brightness(200),
-         brightnessCmd('/camera_control.cgi?param=1&value='),
-         contrast(4), 
-         contrastCmd('/camera_control.cgi?param=2&value='),
-         picCmd('/snapshot.cgi?resolution=32&user=scrapsec&pwd=lakewould'),
-         screen(50, 74, point(200,4)),
-         imageSize(580,440),
-	 layout([
-		 cellstat(cellstat,below,[od(0.4),temp(37.0),shape(20,9),CF]),
-		 spacer(     x1, next_row, [color(blue)]),
-		 snapshot(  cam, next_row, [ image('mypic1.jpg'),shape(30,30)]),
-		 spacer(     x2, next_row, []),
-		 lagoon( lagoon1, next_row, [temp(35.0), LS, SF]),
-		 lagoon( lagoon2, right,    [temp(35.0), LS, SF]),
-		 lagoon( lagoon3, right,    [temp(34.5) ,LS, SF]),
-		 spacer(      x3, next_row, [color(darkgreen)]),
-		 sampler(autosampler, next_row, [shape(40,8),SF])
-                ])
-	 ]) :-
- LS = shape(24,8),
-% LF = font(font(times,roman,14)),
- CF = font(font(times,roman,18)),
- SF = font(font(times,roman,20)).
-
-
-
[…]

### 2020-03-05 host/param.h

@@ -19,11 +19,11 @@
 #define LED             13    // Arduino LED
 #define LASER           11    // Turbidity LASER (PWM on Nano or Mega)
 #define JARLIGHT        7     // Meniscus light
-#define NUTRIENT        4     // Only one valve for Cellstat
-#define INDUCE1         6     // Only one valve for Cellstat
-#define INDUCE2         8     // Only one valve for Cellstat
+#define NUTRIENT        4     // V0 is Nutrient supply
+#define INDUCE1         5     // 
+#define INDUCE2         8     // 
 #define HOSTOUT         5     // Output to Waste
-#define MIXER           3     // PWM for 12V motor
+#define MIXER           3     // PWM for 5V motor
 
 #define DEFAULT_CYCLETIME       20  // Seconds
 

### 2020-03-05 host/pid.cpp

@@ -20,7 +20,7 @@
 PID::PID(double* Input, double* Output, double* Setpoint,
         double Kp, double Ki, double Kd, int ControllerDirection)
 {
-	
+    controllerDirection = ControllerDirection;
     myOutput = Output;
     myInput = Input;
     mySetpoint = Setpoint;
@@ -87,7 +87,7 @@ void PID::SetTunings(double Kp, double Ki, double Kd)
    ki = Ki * SampleTimeInSec;
    kd = Kd / SampleTimeInSec;
  
-  if(controllerDirection ==REVERSE)
+  if(controllerDirection == REVERSE)
    {
       kp = (0 - kp);
       ki = (0 - ki);
@@ -166,7 +166,7 @@ void PID::SetControllerDirection(int Direction)
 {
    if(inAuto && Direction !=controllerDirection)
    {
-	  kp = (0 - kp);
+      kp = (0 - kp);
       ki = (0 - ki);
       kd = (0 - kd);
    }   

### 2020-03-05 host/valves.h

@@ -36,7 +36,7 @@ class VALVES
     cycletime = DEFAULT_CYCLETIME*1000; // Cycle in seconds, store as ms
   }
 
-  void setup_valve(int v, byte pn, int tm, byte dir) {
+  void setup_valve(int v, byte pn, uint16_t tm, byte dir) {
     valve_time[v] = tm;
     valve_pin[v] =  pn;
     valve_dir[v] = dir;
@@ -53,7 +53,7 @@ class VALVES
   int getCycletime(void)      { return cycletime/1000;  }
   void setCycletime(int secs) { cycletime = secs*1000; }
 
-  boolean setValve(int pin, int time) {
+  boolean setValve(int pin, uint16_t time) {
     int i;
     for(i=0;i<size;i++) {
       if (valve_pin[i] == pin) {
@@ -112,7 +112,7 @@ boolean checkValves(void) {
     valve_open[v] = 0;
   }
 
-  int *getTimes()            { return &valve_time[0];   }
+  uint16_t *getTimes()       { return &valve_time[0];   }
   int getTime(int vn)        { return valve_time[vn]; }
   int setTime(int vn, int t) { valve_time[vn] = t; }
 
@@ -121,7 +121,7 @@ boolean checkValves(void) {
   byte     flow;
   byte     valve_pin[NUM_VALVES];
   byte     valve_dir[NUM_VALVES];
-  int      valve_time[NUM_VALVES];
+  uint16_t valve_time[NUM_VALVES];
   long int valve_open[NUM_VALVES];  // When was the valve opened
 
   unsigned long lasttime;  // Beginning of current time interval

### 2020-03-05 huxley.pl

@@ -22,9 +22,9 @@ config( [
 
 	 frames(100),       % number of frames for lumosity integration
 	 darkness(60),      % Average pixel threshold to identify darkness
-	 camera(0),
-	 rotate(90),
-	 screen(52,86,point(110,40)),
+	 camera(0),  % value of device number(Windows) or /dev/videoN (Linux)
+	 rotate(90), % Clockwise rotation before image processing
+	 screen(52,86,point(110,40)), % Placement of GUI on desktop
 	 layout([
 		 supply( nutrient, below,  [Supply,levelUnits('L'),v(10)]),
 		 supply( arabinose, right, [Supply,levelUnits(mL)]),

### 2020-03-05 lagoon/param.h

@@ -12,9 +12,10 @@
 
 #define LED            10     // Meniscus light
 #define HEATER         11     // Power resistors
+#define DRIP            2     // Interrupt Drip
 #define MIXER           3     // PWM for 5V motor
 
-#define MIXERSPEED   80      // PWM (5v) value for top mixer speed
+#define MIXERSPEED   120      // PWM (5v) value for top mixer speed
 
 #define DEFAULT_CYCLETIME 30L // Can be changed/stored in EEPROM
 

### 2020-03-05 level.py

@@ -539,10 +539,12 @@ if __name__ == "__main__" :
     global cycletime
     cycletime = 90  # Default
     start = int(time.time())
-#    sys.stdout = os.fdopen(sys.stdout.fileno(), 'w', 0) # No buffering
+    sys.stdout = os.fdopen(sys.stdout.fileno(), 'w', 0) # No buffering
     os.close(sys.stderr.fileno())
-#   sys.stderr = open('/tmp/level.log','a')
-    sys.stderr = open('/home/pi/src/EvoStat/web/level.txt','a')
+    if os.path.exists('/home/pi/src/EvoStat/web/') :
+        sys.stderr = open('/home/pi/src/EvoStat/web/level.txt','a')
+    else :
+        sys.stderr = open('/tmp/level.log','a')
     memuse()
     print("Here I am at new stderr", file=sys.stderr)
     memuse()

### 2020-03-05 module/drip.h

@@ -21,7 +21,7 @@ void reportDRIP(void)
     }
   Serial.println("]).");
 }
-
+// DRIP, pin 2 (param.h)
 void dripCatcher0() // 0 -> 1 transition on pin 2
 {
 long now = millis();
@@ -31,6 +31,7 @@ long now = millis();
      DRIP_time[0] = now;
 }
 
+// DRIP2, pin 3 (param.h)
 void dripCatcher1() // 0 -> 1 transition on pin 6
 {
 long now = millis();

### 2020-03-05 module/param.h

@@ -19,10 +19,10 @@
 #define NUM_VALVES      4
 
 // Only 3 PWM Nano: [3 int],5,6,[9,10 servo],11
-#define DRIP            2    // Interrupt Drip Beam     SC
+#define DRIP            2    // Interrupt Drip Beam     SensorConnector
 #define DRIP2           3    // Aux Interrupt Drip      SC
 #define AIR             4    // AIR PUMP
-#define MENISCUS        5    // Meniscus light     PWM  MC
+#define MENISCUS        5    // Meniscus light     PWM  ModuleConnector
 #define MIXER           6    // PWM for 5V motor   PWM  MC
 #define WASTE           7    // Output to Waste (12V Solenoid Valve)
 #define VALVEPIN        8    // Control of valve servo

### 2020-03-05 module/valves.h

@@ -16,6 +16,10 @@
 //                            close valves after their "open time" 
 // void openValve(int v)    - Open the valve for its "open time" duration
 //
+//  Extend the notion of valve to include a heat and air valves to
+// control the "heat on" and "air on" intervals as part of the total cycle.
+//
+
 
 //#include "Arduino.h"
 //#include "param.h" // No need since this file is included

### 2020-03-05 nextlevel.py

@@ -68,7 +68,7 @@ def exportImage(image) :
         (y1,x1,y2,x2) = params['lagoonRegion']
         cv2.rectangle(image,(x1,y1),(x2,y2),(250,250,0),2)
         for i in range(4):
-            ((lx1,ly1),(lx2,ly2)) = lagoon_level_bbox_xy(i)
+            ((lx1,ly1),(lx2,ly2)) = lagoon_level_bbox_thirds(i)
             cv2.rectangle(image,(lx1,ly1),(lx2,ly2),
                           (100, 255-i*60, i*60), 2)
         name = "./web/phagestat.jpg"
@@ -153,8 +153,8 @@ def showCoord(img,pt,color,size) :
 
 def showLevel(img, bb, lvl, color) :  # Return % from bootom
     height = bb[2]-bb[0]
-    w2 = (bb[3]-bb[1])/2
-    (px,py) = ( bb[3], bb[0]+lvl )
+    w2 = int( (bb[3]-bb[1])/2 )
+    (px,py) = ( int(bb[3]), int(bb[0]+lvl) )
     if (lvl == 0) :
         pc = 100
     else :
@@ -246,6 +246,17 @@ def lagoon_level_bbox_xy(lnum) :
     right = x1 + w16 + w8 - 8
     return ((left+(lnum*w4), y1+8),(right+(lnum*w4),y2-8))
 
+def lagoon_level_bbox_thirds(lnum) :
+    global params
+    (y1,x1,y2,x2) = params['lagoonRegion']
+    w3 = int((x2-x1)/3)
+    w6 = w3/1.5
+    w12 = w6/3
+    left = x1 + w12 + 8
+    right = x1 + w12 + w6 - 8
+    return ( ( int(left+(lnum*w3)), int(y1+8) ),
+             ( int(right+(lnum*w3)), int(y2-8) ) )
+
 if __name__ == "__main__" :
     plog("openCV('" + str(cv2.__version__) + "').")
     params = settings()
@@ -270,7 +281,7 @@ if __name__ == "__main__" :
     llist = [ showLevel(img, brect, lev, (0,255,255)) ]
     img2 = img
     for i in range(params['numLagoons']) :
-        ( (lx1,ly1), (lx2,ly2) ) = lagoon_level_bbox_xy(i)
+        ( (lx1,ly1), (lx2,ly2) ) = lagoon_level_bbox_thirds(i)
         lbb = (ly1,lx1,ly2,lx2)
         llev = 0
         tries = 5

### 2020-03-05 perpetua.sh

@@ -1,10 +1,11 @@
 #!/bin/bash
-cd /home/peter/src/EvoStat
+cd ~/src/EvoStat
 while true; do
     EPID=`pgrep evostat`
     if [ "$EPID" = "" ]; then
+	mv web/evostat.txt evostat.txt.`date +"%d-%m-%Y-%H:%M:%S"`
 	echo "EvoStat is down. Launching new Evostat" >>evostat.log
-	nohup ./evostat 2>&1 >>evostat.log &
+	nohup ./evostat 2>&1 >evostat.txt &
     fi
     echo "EvoStat PID is [" $EPID "]"
     sleep 5

### 2020-03-05 suppress_stdout_stderr.py

@@ -1,22 +1,13 @@
-# Define a context manager to suppress stdout and stderr.
 import os
 class suppress_stdout_stderr(object):
     '''
-    A context manager for doing a "deep suppression" of stdout and stderr in 
-    Python, i.e. will suppress all print, even if the print originates in a 
-    compiled C/Fortran sub-function.
-       This will not suppress raised exceptions, since exceptions are printed
-    to stderr just before a script exits, and after the context manager has
-    exited (at least, I think that is why it lets exceptions through).      
-    To use this say:
-
+    Context manager for "deep suppression" of stdout and stderr
     with suppress_stdout_stderr():
         rogue_function()
     '''
     def __init__(self):
-        # Open a pair of null files
+        # Open null files and save stdout(1) and stderr(2)
         self.null_fds =  [os.open(os.devnull,os.O_RDWR) for x in range(2)]
-        # Save the actual stdout (1) and stderr (2) file descriptors.
         self.save_fds = (os.dup(1), os.dup(2))
 
     def __enter__(self):
@@ -25,9 +16,8 @@ class suppress_stdout_stderr(object):
         os.dup2(self.null_fds[1],2)
 
     def __exit__(self, *_):
-        # Re-assign the real stdout/stderr back to (1) and (2)
+        # Re-assign stdout/stderr back to (1) and (2) and close null files
         os.dup2(self.save_fds[0],1)
         os.dup2(self.save_fds[1],2)
-        # Close the null files
         os.close(self.null_fds[0])
         os.close(self.null_fds[1])

### 2020-03-05 web/controlpathe.pl

@@ -17,7 +17,9 @@ space(_) --> [].
 inputItem(Type) --> [ Name ], var_label(P),
 		    { component(Name,Type,Obj),
 		      concat_atom([Name,'_',P],Var),
-                      webValue(Obj,P,Val)           },
+                      webValue(Obj,P,Val)
+%		      (Val = 0 -> DVal = 'N/A' ; DVal=Val)
+		    },
 		    [ input([type=text,name=Var,value=Val]) ],
 		    var_units(P), nl, space(Type).
 
@@ -33,10 +35,9 @@ controlpathe(_Req) :-
     flatten([ input([type=text,id=motd,name=motd,value=Message]),br([]),
 	      Fields, input([type=submit,name=submit,value='Submit']),
 	input([ type=button,name='Cancel',value='Cancel',
-        onClick='window.location="/web/pathe.pl"'])],Inputs),
-    reply_html_page([title('Pathe Control Panel'),
-	             script([ language(javascript) ],[])],
-	body(background('/web/images/brushed.jpg'),
+		onClick='window.location="/web/pathe.pl"'])],Inputs),
+    defaultHead('Pathe Control Panel',Head),
+    reply_html_page(Head, body(background('/web/images/brushed.jpg'),
 	     form(action='./change.pl', Inputs))),
     retract(html_syntax).
 

### 2020-03-05 web/f.m

@@ -1,44 +1,0 @@
-% LAGOON MODEL:  FOUR COUPLED DIFFERENTIAL EQUATIONS:
-%
-% UNINFECTED  -> ADSORBED  -> PRODUCTIVE  ->  PHAGE
-%  (cells)        (cells)      (cells)      (particles)
-
-function xdot = f(x,t)
-  global kg ad ec pp fr h0 vol;
-  pm = 1; % This is what makes the g-model different (from f)
-
-  lkg = delta(kg, 0.02*pm); % Local (varied) normal E. coli growth constant
-  lad = delta(ad, 0.1*pm);  % Local (varied) adsorption coefficient
-  lec = delta(ec, 0.1*pm);  % Local (varied) rate exiting eclipse
-  lpp = delta(pp, 0.05*pm);  % Local (varied) phage production per cell
-  dc = fr; % new variable so we could model dynamic modification
-
-% Inhibited growth rates for subsequent stages of E. coli
-% Ultimately this should be measured for given strains of E. coli and M13
-  inhibit = 0.9;
-% Infected cells have inhibited growth rate relative to healthy host cells
-  kgp = lkg*inhibit;
-% Productive cells have inhibited growth rate relative to merely infected cells
-  kgpp = kgp*inhibit;
-
-% INFLOW = Flow rate in volumes/hour * concentration cells/ML * volume mL = number of cells
-% Growth = Number of Cells * Growth constant (doubling time?)             = number of cells
-% Infection = (Total Number of Cells * Total Number of Phage * Adsorption Constant)/total Volume = Fraction Infected
-  
-% (1) Uninfected Host flowing in at Dilution Constant * Cells/mL
-%           + INFLOW       +   GROWTH   - OUTFLOW  - TRANSFORMATION
-  xdot(1) =   dc*h0*vol    +  lkg*x(1)  - dc*x(1)  - (lad*x(1)*x(4))/vol;
-
-% (2) Infected Host (previous TRANSFORMED is this equation's INFLOW)
-  xdot(2) =  (lad*x(1)*x(4))/vol +  kgp*x(2)  - dc*x(2)  -  lec*x(2);
-
-% (3) Productive Host: INFLOW from Eclipse state, plus growth, minus dilution 
-
-  xdot(3) =  lec*x(2)      + kgpp*x(3)  - dc*x(3);  % No transformation output
-  
[…]

### 2020-03-05 web/g.m

@@ -1,43 +1,0 @@
-% LAGOON MODEL:  FOUR COUPLED DIFFERENTIAL EQUATIONS:
-%
-% UNINFECTED  -> ADSORBED  -> PRODUCTIVE  ->  PHAGE
-%  (cells)        (cells)      (cells)      (particles)
-
-function xdot = g(x,t)
-  global kg ad ec pp fr h0 vol;
-  pm = -1; % This is what makes the g-model different (from f)
-
-  lkg = delta(kg, 0.02*pm); % Local (varied) normal E. coli growth constant
-  lad = delta(ad, 0.1*pm);  % Local (varied) adsorption coefficient
-  lec = delta(ec, 0.1*pm);  % Local (varied) rate exiting eclipse
-  lpp = delta(pp, 0.05*pm);  % Local (varied) phage production per cell
-  dc = fr; % new variable so we could model dynamic modification
-
-% Inhibited growth rates for subsequent stages of E. coli
-% Ultimately this should be measured for given strains of E. coli and M13
-  inhibit = 0.9;
-% Infected cells have inhibited growth rate relative to healthy host cells
-  kgp = lkg*inhibit;
-% Productive cells have inhibited growth rate relative to merely infected cells
-  kgpp = kgp*inhibit;
-
-% INFLOW = Flow rate in volumes/hour * concentration cells/ML * volume mL = number of cells
-% Growth = Number of Cells * Growth constant (doubling time?)             = number of cells
-% Infection = (Total Number of Cells * Total Number of Phage * Adsorption Constant)/total Volume = Fraction Infected
-  
-% (1) Uninfected Host flowing in at Dilution Constant * Cells/mL
-%           + INFLOW       +   GROWTH   - OUTFLOW  - TRANSFORMATION
-  xdot(1) =   dc*h0*vol    +  lkg*x(1)  - dc*x(1)  - (lad*x(1)*x(4))/vol;
-
-% (2) Infected Host (previous TRANSFORMED is this equation's INFLOW)
-  xdot(2) =  (lad*x(1)*x(4))/vol +  kgp*x(2)  - dc*x(2)  -  lec*x(2);
-
-% (3) Productive Host: INFLOW from Eclipse state, plus growth, minus dilution 
-
-  xdot(3) =  lec*x(2)      + kgpp*x(3)  - dc*x(3);  % No transformation output
-  
[…]

### 2020-03-05 web/graph.pl

@@ -38,5 +38,18 @@ test :-
     consult('datalog.txt'),
     setof(Time:Data, toCSV(Time,Data), AllData),
     open(httltloo,write,F),
+    format(F,'htemp,turbidity,hlevel,ltemp,llevel,v0,v1\n',[]),
     data_sequence(AllData,F),
     close(F).
+
+range(What,Who,Min,Max) :-
+    member(What,[temperature,turbidity,level,hostValve,lagoonValve]),
+    Term =.. [What, Who, Time, Val],
+    setof(range(Val,What),What^Who^Time^Val^Term,Solns),
+    Solns = [MinTerm|_],
+    append(_,[MaxTerm],Solns),
+    arg(1,MinTerm,Min),
+    arg(1,MaxTerm,Max).
+    
+    
+% average(Var,Goal,-Average)

### 2020-03-05 web/phagepop.pl

@@ -105,8 +105,8 @@ phagepop(_Req) :-
 	br([],[]),
 	pre([],[' ']),
 	br([],[]),
-	font([size('-2')],
-	['Be patient, E.g. 300 hour simulation requires more than 1 minute'])],
+	font([size('+1')],
+	['Be patient, E.g. 300 hour simulation requires more than 1 minute on 2.1Ghz CPU'])],
         Inputs),
     defaultHead('Pathe Population Model',Head),
     reply_html_page(Head,

### 2020-03-05 web/phagestat.pl

@@ -2,14 +2,14 @@
 % Problems with image cacheing, we always want to see the latest
 phagestat(_) :-
   nocacheHead('Inside Phagestat',NCH),
-  reply_html_page(default, NCH,
+  reply_html_page(default,
+		  [meta(['http-equiv'(refresh),content(5)],[])|NCH],
 	  body(style('background-color:grey'),
-	       [pre([style='top-margin:10%'],[' ']),
-		center(img([src('./phagestat.jpg'),width('96%')],[])),
-		video([width(320),height(240),autoplay],
-		   source([src('./timelapse.avi'),type('video/x-msvideo')],[]))
+	       [pre([style='top-margin:5%'],[' ']),
+		center(img([src('./phagestat.jpg'),width('90%')],[]))
 	       ])).
 
 phagestat(Request) :-
     errorPage(Request, 'Error creating PhageStat image').
+
 

### 2020-03-05 web/plot.pl

@@ -1,7 +1,7 @@
 plot(_Req) :-
-    defaultHead('Level/Temperature Plot', Head),
-    shell('web/ptemp.py flog.txt'),
-    reply_html_page(Head,body(background('plot.png'),[])),
+    defaultHead('Population Plot', Head),
+%   shell('web/ptemp.py flog.txt'),
+    reply_html_page(Head,body(background('phagepop.png'),[])),
     !. % det only necessary for interactive debugging
 
 plot(Request) :- errorPage(Request,'EvoStat Plot page error').

### 2020-03-05 web/run_model.pl

@@ -1,5 +1,9 @@
 :- use_module(library(process),[process_create/3]).
 
+% Simulated input parameter to debug this web page
+debug_run_model(search, [kg='60',ad='3.0e-10',ec='6',pp='2',p0='10000000',fr='1',
+			 h0='100000000000',vol='40',dur='40',lsl='0',bl='1.0e3',submit='Submit']).
+
 octaveParameters([])            --> [].
 octaveParameters([X=Y|Ps])      -->
     { atom_concat('.',_,Y)->atomic_list_concat(['0',Y],NY);Y=NY},
@@ -8,6 +12,9 @@ octaveParameters([X=Y|Ps])      -->
 
 run_model(Req) :-
     memberchk(search(S),Req),
+    open('./web/rm_search.txt',write,STxt),
+    writeq(STxt,S),
+    close(STxt),
     select(submit=_, S,S1),
     octaveParameters(S1,Tokens,[]),
     flatten(Tokens,List),
@@ -28,7 +35,4 @@ run_model(Req) :-
 
 run_model(Request) :-
       errorPage(Request, 'Error changing parameters').
-
-
-
 

### 2020-03-05 web/templot.py

@@ -36,8 +36,8 @@ ax.tick_params(which='minor', length=5, labelsize=3, labelcolor='0.25')
 
 ax.grid(linestyle="--", linewidth=0.5, color='.25', zorder=-10)
 
-ax.plot(X/86400.0, Y2, c=(1.00, 0.25, 0.25), lw=0.2, label="Lagoon", zorder=11)
-ax.plot(X/86400.0, Y1, c=(0.25, 0.25, 1.00), lw=0.2, label="Host Culture", zorder=10)
+ax.plot(X/86400.0, Y2, c=(1.00, 0.25, 0.25), lw=0.3, label="Lagoon", zorder=11)
+ax.plot(X/86400.0, Y1, c=(0.25, 0.25, 1.00), lw=0.3, label="Host Culture", zorder=10)
 
 ax.set_title("EvoStat Temperature January 7-26, 2019", fontsize=20, verticalalignment='bottom')
 ax.set_xlabel("Time (days)")
